---
title: "Spatial `K`luster Analysis by Tree Edge Removal"
subtitle: "Experiment for Randomness"
author: "Maxwell"
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: show
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

# Preparations {.tabset .tabset-fade .tabset-pills}

## Load libraries
```{r, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(DT)
library(gt)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(RColorBrewer)  # display.brewer.all()
library(extrafont)  # "Candara"
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping: skater
library(rmapshaper)  # helpful functions for mapping
library(tmap)  # map viewer
library(leaflet)  # map viewer
library(mapview)  # map viewer

# Descriptive Statistics
library(naniar)
library(psych)
library(gtsummary)
```

## Environment
```{r Environment}
# rgb(73, 81, 112, maxColorValue = 255): "#495170"
col.os <- "#414a4c"
col.cb <- "#b0b7c6"
col.rm <- "#e3256b"
theme_set(theme_minimal(base_family = "Candara"))
options(dplyr.summarise.inform = TRUE)
```

# Load DataSets {.tabset .tabset-fade .tabset-pills}
```{r Load DataSets}
bh <- sf::read_sf(
    system.file("etc/shapes/bhicv.shp", package = "spdep"), 
    quiet = TRUE
)
print(st_crs(bh))
st_crs(bh) <- "+proj=longlat +ellps=WGS84"  # set CRS
```


# Preprocessing for skater algorithm {.tabset .tabset-fade .tabset-pills}
## Standardize Node 4 Features
```{r}
bh.index.scale <- scale(
    x = bh %>% as.data.frame() %>% select(HLCI, ELCI, CLCI, ELCI_1)
)
bh.index.scale.df <- bh.index.scale %>% as.data.frame()
```

## Compute Neighboorhod list with `spdep::poly2nb`
```{r}
bh.nb <- spdep::poly2nb(bh)
spdep::card(bh.nb)
```

## Computing costs with `spdep::nbcosts`
```{r}
bh.costs <- spdep::nbcosts(
    nb = bh.nb, 
    data = bh.index.scale,
    method = c("euclidean", "maximum", "manhattan", "canberra", 
               "binary", "minkowski", "mahalanobis")
)
cat(class(bh.costs))
```

## Making listw with `spdep::nb2listw`
```{r}
unlist(bh.costs)
bh.listw <- spdep::nb2listw(neighbours = bh.nb, bh.costs, style = "B")
bh.listw
```

## Find a Minimum Spanning Tree with `spdep::mstree` and Plot
```{r, fig.width=10, fig.height=10}
for (i in seq(1, nrow(bh), 10)) {
    bh.mst <- spdep::mstree(nbw = bh.listw, ini = i)
    if (i == 1) {
        plot(sf::st_geometry(bh), border = col.os, add = FALSE)
        plot(bh.mst, coordinates(as(bh, "Spatial")), 
             cex.circles = 0., cex.labels = 0.5, 
             fg = col.os, add = TRUE, col = col.rm)
    } else {
        plot(bh.mst, coordinates(as(bh, "Spatial")), 
             cex.circles = 0., cex.labels = 0.5, 
             fg = col.os, add = TRUE, col = col.rm)
    }
}
```


# Regionalization with `spdep::skater` {.tabset .tabset-fade .tabset-pills}
## ncuts = 2 (ncluster = 3)
```{r}
bh.reg.1 <- spdep::skater(
    edges = bh.mst[, 1:2], 
    data = bh.index.scale.df,
    ncuts = 2
)
summary(bh.reg.1)
```

```{r}
cat(names(bh.reg.1), sep = "\n")
table(bh.reg.1$groups)
```

```{r, fig.height=10}
bh %>% bind_cols(
    tibble(GROUP = factor(bh.reg.1$groups, levels = c(1, 2, 3)))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "Region: ", palette = "Vivid") + 
    theme(legend.position = "none")
```
