---
title: "Spatial `K`luster Analysis by Tree Edge Removal"
subtitle: ""
author: "Maxwell"
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: show
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

Tutorial for regionalization function `skater`.

References:  

1. AssunÇão RM, Neves MC, Câmara G, Da Costa Freitas C. Efficient regionalization techniques for socio‐economic geographical units using minimum spanning trees. International Journal of Geographical Information Science. 2006;20: 797–811. doi:10.1080/13658810600665111

2. https://r-spatial.github.io/spdep/reference/skater.html

# Preparations {.tabset .tabset-fade .tabset-pills}

## Load libraries
```{r, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(DT)
library(gt)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(RColorBrewer)  # display.brewer.all()
library(extrafont)  # "Candara"

# Geospatial Analysis
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping: skater
library(rmapshaper)  # helpful functions for mapping
library(tmap)  # map viewer
library(leaflet)  # map viewer
library(mapview)  # map viewer

# Descriptive Statistics
library(naniar)
library(psych)
library(gtsummary)
```

## Environment
```{r Environment}
# rgb(73, 81, 112, maxColorValue = 255): "#495170"
col.os <- "#414a4c"
col.cb <- "#b0b7c6"
col.rm <- "#e3256b"
theme_set(theme_minimal(base_family = "Candara"))
options(dplyr.summarise.inform = TRUE)
```

# Load DataSets {.tabset .tabset-fade .tabset-pills}

**> ?spdep::bhicv**  
**Data set with 4 life condition indices of Belo Horizonte region**  

The data are collected inthe Atlas of condition indices published by the Joao Pinheiro Foundation and UNDP.

**Format**: A shape polygon object with seven variables  
**id**: The identificator  
**Name**: Name of city  
**Population**: The population of city  
**HLCI**: Health Life Condition Index  
**ELCI**: Education Life Condition Index  
**CLCI**: Children Life Condition Index  
**ELCI_1**: Economic Life Condition Index  

```{r Load DataSets}
bh <- sf::read_sf(
    system.file("etc/shapes/bhicv.shp", package = "spdep"), 
    quiet = TRUE
)
print(st_crs(bh))
st_crs(bh) <- "+proj=longlat +ellps=WGS84"  # set CRS
```
```{r}
ggplot() + 
    geom_sf(mapping = aes(), data = bh,
            color = col.os, fill = col.cb, size = 0.2, alpha = 0.5)
```


# Preprocessing for skater algorithm {.tabset .tabset-fade .tabset-pills}
## Standardize Node 4 Features
```{r}
bh.index.scale <- scale(
    x = bh %>% as.data.frame() %>% select(HLCI, ELCI, CLCI, ELCI_1)
)
bh.index.scale.df <- bh.index.scale %>% as.data.frame()
```

## Compute Neighboorhod list with `spdep::poly2nb`
```{r}
bh.nb <- spdep::poly2nb(bh)
spdep::card(bh.nb)
```

## Computing costs with `spdep::nbcosts`
**method**:  

| "mahalanobis" 
|       or 
| "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowisk"

```{r}
bh.costs <- spdep::nbcosts(
    nb = bh.nb, 
    data = bh.index.scale,
    method = c("euclidean", "maximum", "manhattan", "canberra", 
               "binary", "minkowski", "mahalanobis")
)
cat(class(bh.costs))
```

## Making listw with `spdep::nb2listw`
```{r}
unlist(bh.costs)
bh.listw <- spdep::nb2listw(neighbours = bh.nb, bh.costs, style = "B")
bh.listw
```

## Find a Minimum Spanning Tree with `spdep::mstree` and Plot

**nbw**: An object of listw class returned by nb2listw function. See this help for details.

**ini**: The initial node in the minimal spanning tree. Heuristic

```{r}
bh.mst <- spdep::mstree(nbw = bh.listw, ini = 5)
cat(class(bh.mst))
cat(dim(bh.mst))
bh.mst[1:3, ]
```
```{r, fig.width=10, fig.height=10}
plot(sf::st_geometry(bh), border = col.os)
# ?plot.mst
plot(bh.mst, coordinates(as(bh, "Spatial")), 
     cex.circles = 0., cex.labels = 0.5, fg = col.os, add = TRUE, 
     col = col.rm)
```


# Regionalization with `spdep::skater` {.tabset .tabset-fade .tabset-pills}

**Arguments**  
**edges**: A matrix with 2 colums with each row is an edge  
**data**: A data.frame with data observed over nodes.  
**ncuts**: The number of cuts. `This value specifiedis not the number of clusters, but the number of cuts in the graph, one less than the number of clusters`.  
**crit**: A scalar or two dimensional vector with criteria for groups. Examples: limits of group size or limits of population size. If scalar, is the minimum criteria for groups.  
**vec.crit**: A vector for evaluating criteria.  
**method**: "mahalanobis" or "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowisk"  
**p**: The power of the Minkowski distance.  
**cov**: The covariance matrix used to compute the mahalanobis distance.  
**inverted**: logical. If 'TRUE', 'cov' is supposed to contain the inverse of the covariance matrix.  
  
  
**Value**  
**groups**: A vector with length equal the number of nodes. Each position identifies the group of node.  
**edges.groups**: A list of length equal the number of groups with each element is a set of edges.  
**not.prune**: A vector identifying the groups with are not candidates to partition.  
**candidates**: A vector identifying the groups with are candidates to partition.  
**ssto**: The total dissimilarity in each step of edge removal.

## ncuts = 2 (ncluster = 3)
```{r}
bh.reg.1 <- spdep::skater(
    edges = bh.mst[, 1:2], 
    data = bh.index.scale.df,
    ncuts = 2
)
summary(bh.reg.1)
```

```{r}
cat(names(bh.reg.1), sep = "\n")
table(bh.reg.1$groups)
```

```{r, fig.height=10}
bh %>% bind_cols(
    tibble(GROUP = factor(bh.reg.1$groups, levels = c(1, 2, 3)))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "Region: ", palette = "Vivid") + 
    theme(legend.position = "none")
```


## ncuts = 3 (ncluster = 4)
```{r}
bh.reg.1 <- spdep::skater(
    edges = bh.mst[, 1:2], 
    data = bh.index.scale.df,
    ncuts = 3
)
summary(bh.reg.1)
```

```{r}
table(bh.reg.1$groups)
```

```{r, fig.height=10}
bh %>% bind_cols(
    tibble(GROUP = factor(bh.reg.1$groups, levels = c(1, 2, 3)))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "Region: ", palette = "Vivid") + 
    theme(legend.position = "none")
```

## ncuts = 4 (ncluster = 5)
```{r}
bh.reg.1 <- spdep::skater(
    edges = bh.mst[, 1:2], 
    data = bh.index.scale.df,
    ncuts = 4
)
summary(bh.reg.1)
```

```{r}
table(bh.reg.1$groups)
```

```{r, fig.height=10}
bh %>% bind_cols(
    tibble(GROUP = factor(bh.reg.1$groups, levels = c(1, 2, 3)))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "Region: ", palette = "Vivid") + 
    theme(legend.position = "none")
ggsave("fig/skater_01.jpg", dpi = 300, width = 8, height = 10)
```
