---
title: "spatialsample"
subtitle: "Vignette: 01"
author: "@Maxwell_110"
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 10
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

https://github.com/tidymodels/spatialsample/

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# GeoSpatial Analysis
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping
library(rmapshaper)  # helpful functions for mapping
library(tmap)  # map viewer
library(leaflet)  # map viewer
library(mapview)  # map viewer
library(spatialsample)

# ML
library(tidymodels)
tidymodels::tidymodels_prefer()
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")
# icon.grob <- png_to_grob()
```


# Dataset {.tabset .tabset-fade .tabset-pills}
```{r Dataset}
data("ames", package = "modeldata")
admin.sf <- sf::read_sf(
    "input/Administrative_district_Tokyo_2021/N03-21_13_210101.shp"
) %>% rmapshaper::ms_filter_islands(min_area = 1e8)
```

# Spatial Clustering CV
coords: A vector of variable names, typically spatial coordinates, to partition the data into disjointed sets via k-means clustering. `This argument is ignored (with a warning) if data is an sf object.`
```{r}
cluster.folds <- spatial_clustering_cv(admin.sf)
```

## Get the CV Results
```{r}
admin.sf <- admin.sf %>% add_column(Fold = NA, .after = "N03_007")
for (i in 1:nrow(cluster.folds)) {
    in_id <- cluster.folds$splits[[i]]$in_id
    out_id <- cluster.folds$splits[[i]]$out_id
    admin.sf[out_id, "Fold"] <- paste0("Fold", i)
}
admin.sf <- admin.sf %>% 
    mutate(Fold = factor(Fold, levels = paste0("Fold", 1:10), 
                         labels = paste0("Fold-", 1:10)))
```


# Visuailization
```{r}
autoplot(cluster.folds)
```
```{r fig.height=5}
p1 <- ggplot(data = admin.sf) + 
    geom_sf(aes(fill = Fold), lwd = .1, color = col.tw) + 
    scale_fill_carto_d(palette = "Vivid") + 
    labs(title = "Spatial Clustering CV", 
         subtitle = "Tokyo District") + 
    annotate(geom = "text", x = 139, y = 35.55, label = "@Maxwell_110",
             alpha = 0.2, size = 4, family = "Candara", color = col.tw) +
    theme(
        title = element_text(size = 17, color = col.tw), 
        plot.subtitle = element_text(size = 14, color = col.tw), 
        legend.text = element_text(color = col.tw), 
        legend.position = c(0.98, 0.88), 
        legend.key.size = unit(12, "points"), 
        legend.title = element_blank(), 
        panel.grid = element_blank(), 
        axis.title = element_blank(), 
        axis.text = element_text(color = col.tw, size = 5), 
        plot.background = element_rect(fill = col.os, color = col.os)
    )
plot(p1)
ggsave(filename = "fig/spatialsample_01-1.jpg", 
       width = 10, height = 5, dpi = 200)
```


# References

