---
title: "Tokyo Land Price"
subtitle: "Spatial Clustering"
author: "Maxwell"
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: show
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Preparations {.tabset .tabset-fade .tabset-pills}

## Load libraries
```{r, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(plotly)
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(RColorBrewer)  # display.brewer.all()
library(extrafont)  # "Candara"
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping: skater
library(rmapshaper)  # helpful functions for mapping
library(tmap)  # map viewer
library(leaflet)  # map viewer
library(mapview)  # map viewer

# Descriptive Statistics
library(naniar)
library(psych)
library(gtsummary)

source("utility/helper_functions.R")
```

## Environment
```{r Environment}
col.os <- "#414a4c"  # Outer Space
col.rp <- "#7851a9"  # Royal Purple
col.cb <- "#b0b7c6"  # Cadet Blue
col.el <- "#ceff1d"  # Electric Lime
col.rm <- "#e3256b"  # Razzmatazz
scale.col.1 <- c()
theme_set(theme_minimal(base_family = "Candara"))
options(dplyr.summarise.inform = TRUE)
```

# Load DataSets {.tabset .tabset-fade .tabset-pills}

https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-L01-v3_1.html
  
L01_006/_100: Price (JPY/m$^{2}$)  
L01_007: Price Change (%)  
L01_022: District Code  
L01_023: City  
L01_024: Address  

  
  
```{r Load DataSets}
admin.sf <- sf::read_sf(
    "input/Administrative_district_Tokyo_2021/N03-21_13_210101.shp"
) %>% rmapshaper::ms_filter_islands(min_area = 1e8)

lp.sf <- sf::read_sf("input/Land-Price_Tokyo_2022/L01-22_13.shp")

print(st_crs(admin.sf))  # JGD2011
print(st_crs(lp.sf))  # JGD2000

admin.sf <- admin.sf %>% st_transform("+proj=longlat +ellps=WGS84")
lp.sf <- lp.sf %>% st_transform("+proj=longlat +ellps=WGS84")
```


# Preprocessing {.tabset .tabset-fade .tabset-pills}
## Concat Polygons by District
```{r}
admin.sf <- admin.sf %>% 
  rename(District = N03_007) %>% 
  select(District) %>% 
  group_by(District) %>% 
  summarise(.groups = "drop")
```

## Transform and Rename
```{r}
lp.sf <- lp.sf %>% mutate(
  Price = L01_006 / 1e6, 
  Price_Change = L01_007 / 100, 
) %>% rename(
  District = L01_022, 
  City = L01_023, 
  Address = L01_024
)
```

# Exploratory Data Analysis {.tabset .tabset-fade .tabset-pills}
## Price and Price Change
### Descriptive Statistics
```{r}
lp.sf.stat <- lp.sf %>% select(District, City, Price, Price_Change) %>% 
  group_by(District, City) %>% 
  summarise(
    Price.Mean = round(mean(Price, na.rm = TRUE), 3), 
    Price.Median = round(median(Price, na.rm = TRUE), 3), 
    Price.Std = round(sd(Price, na.rm = TRUE), 3), 
    Price_Change.Mean = round(mean(Price_Change, na.rm = TRUE), 3), 
    Price_Change.Median = round(median(Price_Change, na.rm = TRUE), 3), 
    Price_Change.Std = round(sd(Price_Change, na.rm = TRUE), 3), 
    .groups = "drop"
  ) %>% 
  st_set_geometry(NULL) %>% 
  left_join(admin.sf, by = "District") %>% 
  st_set_geometry(value = "geometry")

view.table(lp.sf.stat %>% st_set_geometry(NULL))
```

### Choropleth Map
```{r, fig.height=10}
p1 <- lp.sf.stat %>% 
  ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf, 
          fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf(mapping = aes(fill = Price.Mean)) + 
  rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
  labs(title = "Mean Price (1e6 JPY)") + 
  theme(title = element_text(size = 15), 
        legend.position = c(0.075, 0.35), 
        plot.margin = unit(c(0, 0, 15, 0), units = "pt")) + 
  guides(colour=guide_colourbar(
    barwidth = 1, barheight = 10, label.position = "right")
  )

p2 <- lp.sf.stat %>% 
  ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf, 
          fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf(mapping = aes(fill = Price.Median)) + 
  rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
  labs(title = "Median Price (1e6 JPY)") + 
  theme(title = element_text(size = 15), 
        legend.position = c(0.075, 0.35), 
        plot.margin = unit(c(15, 0, 0, 0), units = "pt")) + 
  guides(colour=guide_colourbar(
    barwidth = 1, barheight = 10, label.position = "right")
  )

p1 / p2
```

```{r, fig.height=5}
p1 <- lp.sf.stat %>% filter(District <= 13123) %>% 
  ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf %>% filter(District <= 13123), 
          fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf_text(aes(label = City), color = col.os, cex = 2.5) + 
  labs(x = NULL, y = NULL) + 
  theme(plot.margin = unit(c(0, 10, 0, 0), units = "pt"))

p2 <- lp.sf.stat %>% filter(District <= 13123) %>% 
  ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf %>% filter(District <= 13123), 
          fill = NA, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf(mapping = aes(fill = Price.Median), color = "white") + 
  geom_sf_text(aes(label = Price.Median), color = col.os, cex = 2.5) + 
  rcartocolor::scale_fill_carto_c(name = "", palette = "Tropic") + 
  labs(x = NULL, y = NULL) + 
  theme(title = element_text(size = 15), 
        legend.position = c(0.075, 0.35), 
        plot.margin = unit(c(0, 0, 0, 10), units = "pt")) + 
  guides(colour=guide_colourbar(
    barwidth = 1, barheight = 10, label.position = "right")
  )

p1 + p2 + plot_annotation(title = "Median Land Price", subtitle = "(1e6 JPY)")
```

