---
title: "Tokyo Land Price"
subtitle: "Spatial Clustering"
author: "Maxwell"
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: hide
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Preparations {.tabset .tabset-fade .tabset-pills}

## Load libraries
```{r, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(plotly)
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(RColorBrewer)  # display.brewer.all()
library(extrafont)  # "Candara"
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping: skater
library(rmapshaper)  # helpful functions for mapping
library(tmap)  # map viewer
library(leaflet)  # map viewer
library(mapview)  # map viewer

# Descriptive Statistics
library(naniar)
library(psych)
library(gtsummary)

source("utility/helper_functions.R")
```

## Environment
```{r Environment}
col.os <- "#414a4c"  # Outer Space
col.rp <- "#7851a9"  # Royal Purple
col.cb <- "#b0b7c6"  # Cadet Blue
col.el <- "#ceff1d"  # Electric Lime
col.rm <- "#e3256b"  # Razzmatazz
scale.col.1 <- c()
font.1 <- "Candara"
theme_set(theme_minimal(base_family = font.1))
options(dplyr.summarise.inform = TRUE)

t.img <- png::readPNG("fig/twitter.png")
t.grob <- grid::rasterGrob(t.img, 
                           width = unit(0.4, "cm"), 
                           height = unit(0.4, "cm"),
                           interpolate = FALSE)
```

# Load DataSets {.tabset .tabset-fade .tabset-pills}

https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-L01-v3_1.html

Land Price Open Dataset
  
L01_006/_100: Price (JPY/m$^{2}$)  
L01_007: Price Change (%)  
L01_022: District Code  
L01_023: City  
L01_024: Address  
  
  
```{r Load DataSets}
admin.sf <- sf::read_sf(
    "input/Administrative_district_Tokyo_2021/N03-21_13_210101.shp"
) %>% rmapshaper::ms_filter_islands(min_area = 1e8)

lp.sf <- sf::read_sf("input/Land-Price_Tokyo_2022/L01-22_13.shp")

# print(st_crs(admin.sf))  # JGD2011
# print(st_crs(lp.sf))  # JGD2000

admin.sf <- admin.sf %>% st_transform("+proj=longlat +ellps=WGS84")
lp.sf <- lp.sf %>% st_transform("+proj=longlat +ellps=WGS84")
```

```{r}
admin.code <- readxl::read_excel(
  "input/AdminiBoundary_CD.xlsx", skip = 1
) %>% rename("District" = "行政区域コード")
```


# Preprocessing {.tabset .tabset-fade .tabset-pills}
## Concat Polygons by District
```{r}
admin.sf <- admin.sf %>% 
  rename(District = N03_007) %>% 
  select(District) %>% 
  group_by(District) %>% 
  summarise(.groups = "drop") %>% 
  left_join(admin.code %>% select(District, City), by = "District")
```

## Transform and Rename
```{r}
city.rm <- c(
    "Ooshima", 
    "Niijima", 
    "Kouzushima", 
    "Miyake", 
    "Hachijyou", 
    "Ogasawara"
)

lp.sf <- lp.sf %>% mutate(
    Price = L01_006 / 1e6,
    Price_Change = L01_007, 
) %>% rename(
    District = L01_022,
    # City = L01_023,
    Address = L01_024
) %>% 
    left_join(admin.code %>% select(District, City), by = "District") %>% 
    filter(!City %in% city.rm)
```

# Exploratory Data Analysis {.tabset .tabset-fade .tabset-pills}
## Descriptive Statistics
```{r}
lp.sf.stat <- lp.sf %>% select(District, City, Price, Price_Change) %>% 
  group_by(District, City) %>% 
  summarise(
    Price.Mean = round(mean(Price, na.rm = TRUE), 3), 
    Price.Median = round(median(Price, na.rm = TRUE), 3), 
    Price.Std = round(sd(Price, na.rm = TRUE), 3), 
    Price_Change.Mean = round(mean(Price_Change, na.rm = TRUE), 3), 
    Price_Change.Median = round(median(Price_Change, na.rm = TRUE), 3), 
    Price_Change.Std = round(sd(Price_Change, na.rm = TRUE), 3), 
    .groups = "drop"
  ) %>% 
  st_set_geometry(NULL) %>% 
  left_join(admin.sf %>% select(District), by = "District") %>% 
  st_set_geometry(value = "geometry")

view.table(lp.sf.stat %>% st_set_geometry(NULL))
```

## Price and Price_Change Distribution
```{r, fig.height=5}
p1 <- lp.sf %>% select(Price) %>% 
  ggplot() + 
  geom_histogram(mapping = aes(x = Price), binwidth = 0.2, 
                 fill = col.cb, color = "white") + 
  xlim(NA, 20) + 
  labs(x = NULL, y = NULL, title = expression("Price (1e6 JPY)"))

p2 <- lp.sf %>% select(Price_Change) %>% 
  ggplot() + 
  geom_histogram(mapping = aes(x = Price_Change), bins = 100, 
                 fill = col.cb, color = "white") + 
  labs(x = NULL, y = NULL, title = expression("Price Change (%)"))

p1 / p2
```

```{r}
# boxplot
```


## Choropleth Map: Tokyo
```{r, fig.height=15}
p1 <- lp.sf.stat %>% 
    ggplot() + 
    geom_sf(mapping = aes(), data = admin.sf, 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price.Mean), color = "white", 
                 cex = 2.5, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "Mean Price (1e6 JPY)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 30, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 10, label.position = "right", 
        ticks.linewidth = 3
    ))

p2 <- lp.sf.stat %>% 
    ggplot() + 
    geom_sf(mapping = aes(), data = admin.sf, 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Median)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price.Median), color = "white", 
                 cex = 2.5, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "Median Price (1e6 JPY)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 30, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 10, label.position = "right", 
        ticks.linewidth = 3
    ))

p3 <- lp.sf.stat %>% 
    ggplot() + 
    geom_sf(mapping = aes(), data = admin.sf, 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price_Change.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "ArmyRose") + 
    geom_sf_text(aes(label = Price_Change.Mean), color = "white", 
                 cex = 2.5, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "Mean Price Change (%)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 10, label.position = "right", 
        ticks.linewidth = 3
    ))

p1 / p2 / p3
ggsave("fig/Land-price_Tokyo.jpg", dpi = 300, width = 10, height = 15)
```

## Choropleth Map: Central Tokyo
```{r, fig.height=24, fig.width=8}
p1 <- lp.sf.stat %>% filter(District <= 13123) %>% 
    ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf %>% filter(District <= 13123), 
          fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf_text(aes(label = City), color = col.os, 
               cex = 4, family = "Candara") + 
  labs(x = NULL, y = NULL, title = "Central Tokyo Districts") + 
  theme(title = element_text(size = 15), 
        plot.margin = unit(c(0, 0, 40, 0), units = "pt"))

p2 <- lp.sf.stat %>% filter(District <= 13123) %>% 
  ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf %>% filter(District <= 13123), 
          fill = NA, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf(mapping = aes(fill = Price.Median), color = "white") + 
  geom_sf_text(aes(label = Price.Median), color = col.os, 
               cex = 4, family = "Candara") + 
  rcartocolor::scale_fill_carto_c(name = "", palette = "Tropic") + 
  labs(x = NULL, y = NULL, title = "Median Land Price (1e6 JPY)") + 
  theme(title = element_text(size = 15), 
        legend.position = c(0.05, 0.3), 
        plot.margin = unit(c(0, 0, 40, 0), units = "pt")) + 
  guides(fill = guide_colourbar(
    barwidth = 0.5, barheight = 15, label.position = "right", 
    ticks.linewidth = 3
  ))

p3 <- lp.sf.stat %>% filter(District <= 13123) %>% 
  ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf %>% filter(District <= 13123), 
          fill = NA, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf(mapping = aes(fill = Price_Change.Mean), color = "white") + 
  geom_sf_text(aes(label = Price_Change.Mean), color = col.os, 
               cex = 4, family = "Candara") + 
  rcartocolor::scale_fill_carto_c(name = "", palette = "ArmyRose") + 
  labs(x = NULL, y = NULL, title = "Mean Land Price Change (%)") + 
  theme(title = element_text(size = 15), 
        legend.position = c(0.05, 0.3), 
        plot.margin = unit(c(0, 0, 0, 0), units = "pt")) + 
  guides(fill = guide_colourbar(
    barwidth = 0.5, barheight = 15, label.position = "right", 
    ticks.linewidth = 3
  ))

p1 / p2 / p3
ggsave("fig/Land-price_Central-Tokyo.jpg", dpi = 300, width = 10, height = 30)
```


# Spatial Clustering with SKATER
## Standardize Node Features
```{r}
lp.stat.scale <- lp.sf.stat %>% 
    as.data.frame() %>% 
    select(-District, -City, -Price_Change.Median, -geometry) %>% 
    scale() %>% as.data.frame()
```

## Compute Neighboorhod list with `spdep::poly2nb`
Works fine with `MULTIPOLYGON`, but not with NA.
```{r}
lp.stat.nb <- spdep::poly2nb(lp.sf.stat %>% select(geometry))
spdep::card(lp.stat.nb)
```

## Compute costs with `spdep::nbcosts`
```{r}
lp.stat.costs <- spdep::nbcosts(
    nb = lp.stat.nb, 
    data = lp.stat.scale,
    method = c("euclidean", "maximum", "manhattan", "canberra", 
               "binary", "minkowski", "mahalanobis")
)
# cat(class(lp.stat.costs))
```

## Making listw with `spdep::nb2listw`
```{r}
# unlist(lp.stat.costs)
lp.stat.listw <- spdep::nb2listw(neighbours = lp.stat.nb, 
                                 glist = lp.stat.costs, style = "B")
lp.stat.listw
```

## Find a Minimum Spanning Tree with `spdep::mstree` and Plot
```{r}
lp.stat.mst <- spdep::mstree(nbw = lp.stat.listw, ini = 5)
# cat(class(lp.stat.mst))
cat(dim(lp.stat.mst))
lp.stat.mst[1:3, ]
```
```{r, fig.width=10, fig.height=6.5}
plot(sf::st_geometry(lp.sf.stat), border = col.os)
# ?plot.mst
plot(lp.stat.mst, coordinates(as(lp.sf.stat, "Spatial")), 
     cex.circles = 0., cex.labels = 0.5, fg = col.os, add = TRUE, 
     col = col.rm)
```

## Regionalization {.tabset .tabset-fade .tabset-pills}
### ncuts = 2 (ncluster = 3)
```{r}
lp.stat.reg.2 <- spdep::skater(
    edges = lp.stat.mst[, 1:2], 
    data = lp.stat.scale,
    ncuts = 2
)
summary(lp.stat.reg.2)
# cat(names(lp.stat.reg.2), sep = "\n")
table(lp.stat.reg.2$groups)
```

```{r, fig.height=20}
p1 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Median)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price.Median), color = "white",
                 cex = 3, family = "Candara") +
    labs(x = NULL, y = NULL, title = "Median Price (1e6 JPY)") + 
    theme(title = element_text(size = 15),
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p2 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price_Change.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price_Change.Mean), color = "white",
                 cex = 3, family = "Candara") +
    labs(x = NULL, y = NULL, title = "Mean Price Change (%)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p3 <- lp.sf.stat %>% bind_cols(
    tibble(GROUP = factor(lp.stat.reg.2$groups, levels = 1:3))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 3, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "ncluster = 3") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt"))

p1 / p2 / p3 + 
    plot_annotation(
    title = "Spatial Clustering on Price Stats", 
    subtitle = "", 
    theme = theme(title = element_text(size = 20, family = "Candara"))
    )

ggsave("fig/Spatial-Clustering-Tokyo_n3.jpg", 
       dpi = 300, width = 15, height = 10 * 3)
```

### ncuts = 3 (ncluster = 4)
```{r}
lp.stat.reg.3 <- spdep::skater(
    edges = lp.stat.mst[, 1:2], 
    data = lp.stat.scale,
    ncuts = 3
)
summary(lp.stat.reg.3)
table(lp.stat.reg.3$groups)
```

```{r, fig.height=20}
p1 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Median)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price.Median), color = "white",
                 cex = 3, family = "Candara") +
    labs(x = NULL, y = NULL, title = "Median Price (1e6 JPY)") + 
    theme(title = element_text(size = 15),
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p2 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price_Change.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price_Change.Mean), color = "white",
                 cex = 3, family = "Candara") +
    labs(x = NULL, y = NULL, title = "Mean Price Change (%)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p3 <- lp.sf.stat %>% bind_cols(
    tibble(GROUP = factor(lp.stat.reg.3$groups, levels = 1:4))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 3, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "ncluster = 4") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt"))

p1 / p2 / p3 + 
    plot_annotation(
    title = "Spatial Clustering on Price Stats", 
    subtitle = "", 
    theme = theme(title = element_text(size = 20, family = "Candara"))
    )

ggsave("fig/Spatial-Clustering-Tokyo_n4.jpg", 
       dpi = 300, width = 15, height = 10 * 3)
```

### ncuts = 4 (ncluster = 5)
```{r}
lp.stat.reg.4 <- spdep::skater(
    edges = lp.stat.mst[, 1:2], 
    data = lp.stat.scale,
    ncuts = 4
)
summary(lp.stat.reg.4)
table(lp.stat.reg.4$groups)
```

```{r, fig.height=20}
p1 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Median)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price.Median), color = "white",
                 cex = 3, family = "Candara") +
    labs(x = NULL, y = NULL, title = "Median Price (1e6 JPY)") + 
    theme(title = element_text(size = 15),
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    )) + 
  annotation_custom(grob = t.grob, 
                    xmin = 139.8, xmax = 139.8, 
                    ymin = 35.83, ymax = 35.83) + 
  annotate(geom = "text", x = 139.83, y = 35.83, 
           label = "@Maxwell_110", size = 3, family = "Candara", color = "gray")

p2 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price_Change.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price_Change.Mean), color = "white",
                 cex = 3, family = "Candara") +
    labs(x = NULL, y = NULL, title = "Mean Price Change (%)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p3 <- lp.sf.stat %>% bind_cols(
    tibble(GROUP = factor(lp.stat.reg.4$groups, levels = 1:5))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 3, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "ncluster = 5") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt"))

p1 / p2 / p3 + 
    plot_annotation(
    title = "Spatial Clustering on Price Stats", 
    subtitle = "", 
    theme = theme(title = element_text(size = 20, family = "Candara"))
    )

ggsave("fig/Spatial-Clustering-Tokyo_n5.jpg", 
       dpi = 300, width = 15, height = 10 * 3)
```

### ncuts = 9 (ncluster = 10)
```{r}
lp.stat.reg.9 <- spdep::skater(
    edges = lp.stat.mst[, 1:2], 
    data = lp.stat.scale,
    ncuts = 9
)
summary(lp.stat.reg.9)
table(lp.stat.reg.9$groups)
```

```{r, fig.height=20}
p1 <- lp.sf.stat %>% bind_cols(
    tibble(GROUP = factor(lp.stat.reg.2$groups, levels = 1:3))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 3, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "ncluster = 3") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 30, 0), units = "pt")) + 
  annotation_custom(grob = t.grob, 
                    xmin = 139.8, xmax = 139.8, 
                    ymin = 35.83, ymax = 35.83) + 
  annotate(geom = "text", x = 139.83, y = 35.83, 
           label = "@Maxwell_110", size = 3, family = "Candara", color = "gray")

p2 <- lp.sf.stat %>% bind_cols(
    tibble(GROUP = factor(lp.stat.reg.4$groups, levels = 1:5))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 3, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "ncluster = 5") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 30, 0), units = "pt"))

p3 <- lp.sf.stat %>% bind_cols(
    tibble(GROUP = factor(lp.stat.reg.9$groups, levels = 1:10))
) %>% ggplot() + 
    geom_sf(mapping = aes(fill = GROUP), 
            size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 3, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "ncluster = 10") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt"))

p1 / p2 / p3 + 
    plot_annotation(
    title = "Spatial Clustering on Price Stats", 
    subtitle = "", 
    theme = theme(title = element_text(size = 20, family = "Candara"))
    )

ggsave("fig/Spatial-Clustering-Tokyo_n3-10.jpg", 
       dpi = 300, width = 15, height = 10 * 3)
```
