---
title: "datasauRus"
subtitle: ""
author: ""
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 10
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

library(datasauRus)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")

col.estyle <- rgb(22, 231, 207, maxColorValue = 255)
# icon.grob <- png_to_grob()
```


# Dataset {.tabset .tabset-fade .tabset-pills}
```{r Dataset}
glimpse(datasaurus_dozen)
```


# Statistics for each dataset
```{r}
datasaurus_dozen %>%
    group_by(dataset) %>%
    summarize(
        mean_x    = mean(x), 
        mean_y    = mean(y), 
        std_dev_x = sd(x), 
        std_dev_y = sd(y), 
        corr_x_y  = cor(x, y), 
        spearman_x_y = cor(x, y, method = "spearman"), 
        kendall_x_y = cor(x, y, method = "kendall")
    )
```


# Comparison plot
```{r}
dataset.sel <- c(
    "away", 
    "circle", 
    "star", 
    "dino"
)

ggplot(datasaurus_dozen %>% filter(dataset %in% dataset.sel), 
       aes(x = x, y = y, colour = dataset)) + 
    geom_point() + 
    theme(legend.position = "none") + 
    facet_wrap( ~ dataset, ncol = 2) + 
    theme(
        axis.title = element_blank(), 
        panel.grid.minor = element_blank()
    )

ggsave("fig/4-types-of-datasets-with-corr0.png", 
       dpi = 300, width = 5, height = 5)
```


# $y = x^3$
```{r}
x <- rnorm(100000)
y <- x ^ 3
# plot(x, y)



ggplot(data = data.frame(x = x, y = y)) + 
    geom_point(aes(x = x, y = y), color = col.estyle) + 
    theme(panel.grid.minor = element_blank())
ggsave("fig/y-x3.png", width = 3, height = 4)

cor(x, y, method = "pearson")
cor(x, y, method = "spearman")
# cor(x, y, method = "kendall")
```


# References
Matejka, J., Fitzmaurice, G., 2017. Same Stats, Different Graphs: Generating Datasets with Varied Appearance and Identical Statistics through Simulated Annealing, in: Proceedings of the 2017 CHI Conference on Human Factors in Computing Systems, CHI ’17. Association for Computing Machinery, New York, NY, USA, pp. 1290–1294. https://doi.org/10.1145/3025453.3025912
