---
title: "lme4 convergence problem"
subtitle: ""
author: "Maxwell"
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: tango
      code_folding: show
      # code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "../..")
```

I sometimes encountered a convergence problem with `lme4` package.
Here I will examine the convergence problem by the following guidance:  
https://m-clark.github.io/posts/2020-03-16-convergence/#remove-zero-random-effects

# Preparations {.tabset .tabset-fade .tabset-pills}

## Load libraries
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"
library(naniar)
library(gtsummary)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(coefplot)
library(mixedup)

source("utility/helper_functions.R")
```

## Environment
```{r Environment}
# rgb(73, 81, 112, maxColorValue = 255): "#495170"
col.1 <- crayons()["Blue Gray"]  # blue
col.2 <- crayons()["Caribbean Green"]  # green
col.3 <- crayons()["Razzmatazz"]  # red
scale.col.1 <- c(col.1, col.3)
font.1 <- "Candara"
theme_set(theme_bw(base_family = font.1))
options(dplyr.summarise.inform = TRUE)
```

# Load DataSets {.tabset .tabset-fade .tabset-pills}
```{r Load DataSets}
df <- readRDS("input/convergence.RDS") %>% 
   mutate(
      SUR.ID = factor(SUR.ID), 
      replicate = factor(replicate),
      Unit = factor(1:n())
   )
```
```{r}
glimpse(df)
view.table(df)
```

# VIF
`GVIF of c.tm.depth is very high.`
```{r}
car::vif(lm(CountDetections ~ . -Unit -Day -FalseDetections, dat = df))
```

# Start from naive GLM
```{r}
glm.1 <- glm(
   cbind(ValidDetections, FalseDetections) ~
      tm:Area + tm:c.distance + c.distance:Area + c.tm.depth:Area +
      c.receiver.depth:Area + c.temp:Area + c.wind:Area + c.tm.depth +
      c.receiver.depth + c.temp + c.wind + tm + c.distance + Area + replicate +
      SUR.ID + Day, 
  data = df, 
  family = binomial(link = "logit")
)

summary(glm.1)
```

- Due to multicollinearity, some coefficients resulted in NA.
```{r}
# glm.2 <- update(glm.1, . ~ . - Day)
glm.2 <- glm(
   cbind(ValidDetections, FalseDetections) ~
      # tm:Area + tm:c.distance + c.distance:Area + #c.tm.depth:Area +
      # c.receiver.depth:Area + c.temp:Area + c.wind:Area + #c.tm.depth +
      c.receiver.depth + c.temp + c.wind + tm + c.distance + Area + replicate + 
      SUR.ID, 
   data = df, 
   family = binomial(link = "logit")
)
summary(glm.2)
```
Now it worked.

# Inital GLME
```{r}
glme.1 <- glmer(
   cbind(ValidDetections, FalseDetections) ~ 
      tm:Area + tm:c.distance + c.distance:Area + c.tm.depth:Area + 
      c.receiver.depth:Area + c.temp:Area + c.wind:Area + c.tm.depth + 
      c.receiver.depth + c.temp + c.wind + tm + c.distance + Area + replicate + 
      (1 | SUR.ID) + (1|Day) + (1|Unit), 
   data = df, 
   family = binomial(link = "logit")
)
```
```{r}
summary(glme.1)
glme.1@optinfo$conv$lme4$messages[[1]]
```
Seems to fail in convergence...

# GLME w/o cross terms
```{r}
glme.2 <- glmer(
   cbind(ValidDetections, FalseDetections) ~
      # tm:Area + tm:c.distance + c.distance:Area + #c.tm.depth:Area +
      # c.receiver.depth:Area + c.temp:Area + c.wind:Area + #c.tm.depth +
      c.receiver.depth + c.temp + c.wind + tm + c.distance + Area + replicate + 
      SUR.ID + 
      (1 | Unit),
   data = df, 
   family = binomial(link = "logit")
)
```
```{r}
summary(glme.2)
glme.2@optinfo$conv$lme4$messages[[1]]
```
Seems to fail to converge without change...

# Check Items
## Singularity
This is a mixed model-specific check, and in general, 
checking singularity goes along with removing zero random effects.  
These days, you will usually get a singularity warning 
when it is likely the case. 
In the past, Bolker suggested checking this problem as follows, 
but for this example concluded the result was not close enough to zero 
to be a real concern. 
The theta below are just our `random effect standard deviations`, 
and I would say that the ones besides Unit probably were not 
meaningfully different from zero.
```{r}
pars <- unlist(getME(glme.1, name = c("theta", "beta")))
print(pars)
```

Nowadays lme4 provides the function `isSingular` which uses the steps above to check the minimum value against some specified tolerance.
```{r}
isSingular(glme.1, tol = 1e-5)
```

## Compare Absolute and Scaled Gradient
For the mixed model setting, Bolker notes the following:

> One general problem is that large scaled gradients are often associated with small absolute gradients: 
we might decide that we are more interested in testing the (parallel) minimum of these two quantities.

We can do this as follows for the initial mixed model.
```{r}
derivs_init <- glme.1@optinfo$derivs
sc_grad_init <- with(derivs_init, solve(Hessian, gradient))

max(abs(sc_grad_init))
max(pmin(abs(sc_grad_init), abs(derivs_init$gradient)))
```

## GLME w/o cross terms and c.distance
```{r}
glme.3 <- glmer(
   cbind(ValidDetections, FalseDetections) ~
      # tm:Area + tm:c.distance + c.distance:Area + #c.tm.depth:Area +
      # c.receiver.depth:Area + c.temp:Area + c.wind:Area + #c.tm.depth +
      c.receiver.depth + c.temp + c.wind + tm + 
      # c.distance + 
      Area + replicate + SUR.ID + (1 | Unit),
   data = df, 
   family = binomial(link = "logit")
)
```
```{r}
summary(glme.3)
glme.3@optinfo$conv$lme4$messages[[1]]
```

# Restart GLME
```{r}
pars2 <- getME(glme.2, name = c("theta", "fixef"))
glme.2.update <- update(object = glme.2, 
                        start = pars2, 
                        control = glmerControl(optCtrl = list(maxfun = 2e4)))
```
```{r}
summary(glme.2.update)
glme.2.update@optinfo$conv$lme4$messages[[1]]
```

# GLME with BOBYQA through all iterations
```{r}
glme.2.update.opt <- update(
   object = glme.2, 
   # start = pars2, 
   control = glmerControl(
      optimizer = "bobyqa", 
      optCtrl = list(maxfun = 2e4)
   ) 
)
```
```{r}
summary(glme.2.update.opt)
glme.2.update.opt@optinfo$conv$lme4$messages[[1]]
```

# Visualization
```{r}
glme.2.df <- glme.2 %>% 
    broom.mixed::tidy(conf.int = T) %>%
    as.data.frame() %>% 
    mutate_if(is.numeric, round, 5) %>% 
    filter(!str_detect(term, "Intercept|sd__"))
glme.2.update.opt.df <- glme.2.update.opt %>% 
    broom.mixed::tidy(conf.int = T) %>%
    as.data.frame() %>% 
    mutate_if(is.numeric, round, 5) %>% 
    filter(!str_detect(term, "Intercept|sd__"))
glm.2.df <- glm.2 %>% 
   broom::tidy() %>% 
   mutate(
      conf.low = estimate - std.error, 
      conf.high = estimate + std.error, 
   ) %>% 
   filter(!str_detect(term, "Intercept|sd__"))
```
```{r, fig.width=12, fig.height=10}
glm.2.p <- glm.2.df %>% 
   ggplot(mapping = aes(x = fct_rev(term), y = estimate, 
                         ymin = conf.low, ymax = conf.high)) + 
   geom_pointrange(color = col.1, size = 0.4) + 
   geom_text(mapping = aes(x = fct_rev(term), 
                           y = estimate,
                           label = round(estimate, 3)), 
             check_overlap = TRUE, size = 3,
             family = "Candara", color = col.1,
             hjust = -0.1, nudge_x = 0.25) + 
   geom_hline(yintercept = 0, color = "slategray", 
              linetype = 2, alpha = 0.5) + 
   lims(y = c(NA, max(glm.2.df$conf.high, na.rm = T) + 0.4)) + 
   labs(x = NULL, y = NULL, title = "GLM w/o level 2") + 
   coord_flip() + 
   theme_minimal() +
   # theme_wsj() + 
   theme(
      title = element_text(family = "Candara", color = col.1, size = 10), 
      axis.text = element_text(family = "Candara", color = col.1, size = 8.5)
   )

glme.2.p <- glme.2.df %>% 
   ggplot(mapping = aes(x = fct_rev(term), y = estimate, 
                         ymin = conf.low, ymax = conf.high)) + 
   geom_pointrange(color = col.2, size = 0.4) + 
   geom_text(mapping = aes(x = fct_rev(term), 
                           y = estimate,
                           label = round(estimate, 3)), 
             check_overlap = TRUE, size = 3,
             family = "Candara", color = col.2,
             hjust = -0.1, nudge_x = 0.25) + 
   geom_hline(yintercept = 0, color = "slategray", 
              linetype = 2, alpha = 0.5) + 
   lims(y = c(NA, max(glme.2.df$conf.high, na.rm = T) + 0.4)) + 
   labs(x = NULL, y = NULL, 
        title = "GLME w/ default opt schedule", 
        subtitle = "BOBYQA => Nelder Mead") + 
   coord_flip() + 
   theme_minimal() +
   theme(
      title = element_text(family = "Candara", color = col.2, size = 10), 
      axis.text = element_text(family = "Candara", color = col.2, size = 8.5)
   )

glme.2.update.opt.p <- glme.2.update.opt.df %>% 
   ggplot(mapping = aes(x = fct_rev(term), y = estimate, 
                         ymin = conf.low, ymax = conf.high)) + 
   geom_pointrange(color = col.3, size = 0.4) + 
   geom_text(mapping = aes(x = fct_rev(term), 
                           y = estimate,
                           label = round(estimate, 3)), 
             check_overlap = TRUE, size = 3,
             family = "Candara", color = col.3,
             hjust = -0.1, nudge_x = 0.25) + 
   geom_hline(yintercept = 0, color = "slategray", 
              linetype = 2, alpha = 0.5) + 
   lims(y = c(NA, max(glme.2.update.opt.df$conf.high, na.rm = T) + 0.4)) + 
   labs(x = NULL, y = NULL, title = "GLME w/ only BOBYQA") + 
   coord_flip() + 
   theme_minimal() +
   theme(
      title = element_text(family = "Candara", color = col.3, size = 10), 
      axis.text = element_text(family = "Candara", color = col.3, size = 8.5)
   )

pw <- glm.2.p + glme.2.p + glme.2.update.opt.p
pw + patchwork::plot_annotation(
   caption = "@Maxwell_110"
)

ggsave("fig/lme4-convergence_01.png", dpi = 300, width = 12, height = 10)
```

