---
title: "GeoStan Vignette"
subtitle: "01:"
author: ""
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: tango
      code_folding: show
      # code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "../..")  # modify with the location of code
```

# Preparations {.tabset .tabset-fade .tabset-pills}

## Load libraries
`Remove library if necessary...`
```{r, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(DT)
library(gt)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"
library(sf)  # map
library(leaflet)  # map
library(mapview)  # map

# Interactivity
library(plotly)
library(crosstalk)

# Parallel
library(purrr)
library(foreach)
library(doParallel)

# Descriptive Statistics
library(naniar)
library(psych)
library(gtsummary)

# Stan
library(geostan)
library(rstan)
library(bayesplot)
library(broom)
library(broom.mixed)

source("utility/helper_functions.R")
```

## Environment
```{r Environment}
# rgb(73, 81, 112, maxColorValue = 255): "#495170"
col.1 <- "#495170"  # blue
col.2 <- "#72faa4"  # green
col.3 <- "#f7be16"  # yellow
col.4 <- "#fa8072"  # red
scale.col.1 <- c(col.1, col.4)
font.1 <- "Candara"
theme_set(theme_bw(base_family = font.1))
options(dplyr.summarise.inform = TRUE)
```

# Load DataSets {.tabset .tabset-fade .tabset-pills}

|  Variable  |  Description  |
| ---- | ---- |
|  GEOID  |  Six digit combined state and county FIPS code  |
|  rate.male  |  Raw (crude) age-specific male mortality rate, 2014-2018  |
|  geometry  |  simple features geometry for county boundaries  |

```{r Load DataSets}
data("georgia")
view.table(georgia)
```

# Spatial EDA
```{r}
sp.diag.eda <- sp_diag(
   y = georgia$rate.male, 
   shape = georgia, 
   name = "mortality rate", 
   plot = FALSE,
   mc_style = c("scatter", "hist"), 
)

p1 <- sp.diag.eda$residual_histogram + 
   theme_minimal(base_size = 7, base_family = font.1)
p2 <- sp.diag.eda$mc_plot + 
   theme_minimal(base_size = 7, base_family = font.1)
p3 <- sp.diag.eda$residual_map + 
   theme_minimal(base_size = 7, base_family = font.1)

p1 + p2 + p3
ggsave("fig/geo_stan_vignette_01-fig1.png", 
       width = 7, height = 2.5, dpi = 200)
```

# CAR model
## Fitting
```{r}
A <- shape2mat(georgia, style = "B")
cars <- prep_car_data(A)
fit <- stan_car(
   deaths.male ~ offset(log(pop.at.risk.male)),
   censor_point = 9,
   data = georgia,
   car_parts = cars,
   family = poisson(),
   cores = 4, # for multi-core processing
   refresh = 0 # to silence some printing
) 
```

## Result visualization
```{r}
sp.diag.car <- sp_diag(
   y = fit, 
   shape = georgia, 
   name = "Residual of deaths", 
   plot = FALSE, 
   w = A
)

p1 <- sp.diag.car$fitted_plot + 
   theme_minimal(base_size = 7, base_family = font.1)
p2 <- sp.diag.car$mc_plot + 
   theme_minimal(base_size = 7, base_family = font.1)
p3 <- sp.diag.car$residual_map + 
   theme_minimal(base_size = 7, base_family = font.1)

p1 + p2 + p3
ggsave("fig/geo_stan_vignette_01-fig2.png", 
       width = 7, height = 2.5, dpi = 200)
```

## Fit summary
```{r}
print(fit)
```

