---
title: "Hospitals in Tokyo"
subtitle: ""
author: "Maxwell"
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
# zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 15
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(pals)
library(broman)  # plot_crayons()
library(ggsci)  # Journal palette
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # fonttable(); "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)

# GeoSpatial Analysis
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping
library(rmapshaper)  # helpful functions for mapping
library(tmap)  # map viewer
library(leaflet)  # map viewer
library(mapview)  # map viewer
library(rnaturalearth)

library(ggmap)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")

fn.prefix <- "_"
# icon.grob <- png_to_grob()

# https://en.wikipedia.org/wiki/World_Geodetic_System#WGS84
crs.base <- "+proj=longlat +datum=WGS84"
register_google(key = readRDS(file = "api/google-api.RDS"))
loc.base <- c(lon = 139.73, lat = 35.70)
```


# Dataset

maptyle:  
- "terrain", "terrain-background", "terrain-labels", "terrain-lines"  
- "satellite", "roadmap", "hybrid"  
- "toner", "toner-2010", "toner-2011", "toner-background", "toner-hybrid", "toner-labels", "toner-lines", "toner-lite"  
- "watercolor"

## ggmap
```{r}
tokyo.s <- get_map(
    location = loc.base, 
    zoom = 12,  # 3 (continent) to 21 (building)
    scale = 3, 
    maptype = "terrain-lines", 
    # source = "google",  # "google", "osm", "stamen", "cloudmademap"
    source = "stamen",  # "google", "osm", "stamen", "cloudmademap"
    color = "color", 
    crop = TRUE
)
ggmap(tokyo.s)
```

## Central Tokyo
```{r}
admin.sf <- sf::read_sf(
    "input/Administrative_district_Tokyo_2021/N03-21_13_210101.shp"
) %>% rmapshaper::ms_filter_islands(min_area = 1e8) %>% 
    st_transform(crs = crs.base)


admin.code <- readxl::read_excel(
    "input/AdminiBoundary_CD.xlsx", skip = 1
) %>% rename("District" = "行政区域コード")

admin.sf <- admin.sf %>% 
    rename(District = N03_007) %>% 
    select(District) %>%
    group_by(District) %>%
    summarise(.groups = "drop") %>%
    left_join(admin.code %>% select(District, City), by = "District") %>% 
    filter(District <= 13123)

admin.df <- fortify(admin.sf)
```

## Hospitals
https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-P04-v3_0.html#prefecture13  
https://nlftp.mlit.go.jp/ksj/gml/codelist/MedClassCd.html  
1: Hospitals  
2: Clinics  
3: Dentists  
```{r}
hos.sf <- sf::read_sf("input/Hospitals_Tokyo_2020/P04-20_13.shp") %>% 
    rename(Hospital.Type = P04_001) %>% 
    filter(Hospital.Type != 3)
```


# Viz {.tabset .tabset-fade .tabset-pills}

```{r}
# Creating north arrow and scale bar, motivation from GISTools package
# https://andrewpwheeler.com/2019/08/07/making-a-hexbin-map-in-ggplot/
arrow_data <- function(xb, yb, len) {
    s <- len
    arrow.x = c(0, 0.5, 1, 0.5, 0) - 0.5
    arrow.y = c(0, 1.7  , 0, 0.5, 0)
    adata <- data.frame(aX = xb + arrow.x * s, aY = yb + arrow.y * s)
    return(adata)
}

scale_data <- function(llx, lly, len, height) {
    box1 <- data.frame(
        x = c(llx, llx + len, llx + len, llx, llx),
        y = c(lly, lly, lly + height, lly + height, lly)
    )
    box2 <- data.frame(
        x = c(llx - len, llx, llx, llx - len, llx - len),
        y = c(lly, lly, lly + height, lly + height, lly)
    )
    return(list(box1, box2))
}

# Example

# x_cent <- 830000
# len_bar <- 3000
# offset_scaleNum <- 64300
# 
# arrow <- arrow_data(xb = x_cent, yb = 67300, len = 2500)
# 
# scale_bxs <- scale_data(
#     llx = x_cent,
#     lly = 65000,
#     len = len_bar,
#     height = 750
# )
# 
# label_data <- data.frame(
#     x = c(x_cent, x_cent - len_bar, x_cent, x_cent + len_bar, x_cent), 
#     y = c(72300, offset_scaleNum, offset_scaleNum, offset_scaleNum, 66500), 
#     lab = c("N", "0", "3", "6", "Kilometers")
# )
# 
# base_map <- ggplot() +
#     geom_polygon(
#         data = PhilBound,
#         size = 1.5,
#         color = 'black',
#         fill = 'darkgrey',
#         aes(x = long, y = lat)
#     ) +
#     geom_polygon(data = arrow, fill = 'black', aes(x = aX, y = aY)) +
#     geom_polygon(data = scale_bxs[[1]],
#                  fill = 'grey',
#                  color = 'black',
#                  aes(x = x,  y = y)) +
#     geom_polygon(data = scale_bxs[[2]],
#                  fill = 'white',
#                  color = 'black',
#                  aes(x = x, y = y)) +
#     geom_text(data = lab_data, size = 4, aes(x = x, y = y, label = lab)) +
#     coord_fixed() + theme_void()
```

## Draw
```{r}
# col.bg <- "gray95"
# col.bg <- "#092532"
col.bg <- "#262A33"
col.fg <- "#393F4D"
arrow <- arrow_data(xb = 139.58, yb = 35.55, len = 0.015)
label.df <- data.frame(x = 139.58, y = 35.582, lab = "N")

p <- ggplot() + 
    geom_sf(data = admin.sf, 
            fill = col.fg, alpha = 0.75, 
            size = 0.3, color = col.bg) + 
    geom_polygon(data = arrow, fill = col.c.b, aes(x = aX, y = aY)) + 
    geom_text(aes(x = x, y = y, label = lab), 
              data = label.df, color = col.c.b, size = 4, 
              family = "Times New Roman") + 
    geom_hex(mapping = aes(x, y), 
             data = hos.sf %>%
                 select(Hospital.Type) %>%
                 sfheaders::sf_to_df(fill = TRUE) %>% 
                 filter(x > 139.55 & y > 35.50), 
             bins = 60, 
             alpha = 0.5, 
             color = col.bg, size = 0.1) + 
    scale_fill_carto_c(direction = -1, palette = "Teal") + 
    labs(
        title = "Hospital Distribution in Central Tokyo", 
        subtitle = " Hexagonal Map", 
        x = NULL, y = NULL, 
        caption = "#30DaysMapChallenge; @Maxwell_110"
    ) + 
    theme(
        title = element_text(
            family = "Times New Roman", face = "bold", 
            color = col.c.b, size = 12
        ), 
        plot.title = element_text(
            margin = margin(t = 15, r = -00, b = -15, l = 00, unit = "pt")
        ), 
        plot.subtitle = element_text(
            family = "Times New Roman", face = "bold", 
            color = col.c.b, size = 10, 
            margin = margin(t = 20, r = -00, b = -20, l = 00, unit = "pt")
        ), 
        plot.caption = element_text(color = col.c.b, size = 6), 
        text = element_text(family = "Times New Roman", color = col.fg), 
        # axis.text = element_text(family = font.base, face = "italic", 
        #                          color = col.c.b, size = 4), 
        axis.text = element_blank(), 
        plot.background = element_rect(fill = col.bg, color = col.bg), 
        # plot includes panel
        # panel.background = element_rect(fill = col.bg, color = col.bg),
        panel.grid = element_blank(), 
        panel.grid.major = element_blank(), 
        strip.text = element_blank(), 
        # legend
        # legend.title = element_blank(),
        # legend.text = element_text(
        #     family = "Times New Roman", face = "plain", 
        #     color = col.c.b, 
        #     size = 5
        # ), 
        legend.position = c(0.875, 0.175), 
        legend.key.size = unit(0.015, "npc")
    ) + 
    guides(fill = guide_colourbar(
        title = "Number of Hospitals", 
        title.theme = element_text(
            family = "Times New Roman", face = "plain", 
            color = col.c.b, size = 5
            # angle = 90
        ), 
        title.position = "bottom", title.hjust = 0, title.vjust = -1, 
        barwidth = 0.4, barheight = 4, 
        label.theme = element_text(
            family = "Times New Roman", face = "plain", 
            color = col.c.b, size = 4
        ), 
        label.position = "right", 
        ticks.linewidth = 3, ticks.colour = col.bg
    ))
p

ggsave("fig/Hospitals-in-Tokyo.jpg", width = 5, height = 5)
```
