---
title: "Tokyo Land Price Viz for 30DaysMapChallenge"
subtitle: "Spatial Clustering with ClustGeo"
author: "Maxwell"
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: hide
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Preparations {.tabset .tabset-fade .tabset-pills}

## Load libraries
```{r, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(plotly)
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(RColorBrewer)  # display.brewer.all()
library(extrafont)  # "Candara"

# GeoSpatial Analysis
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping: skater
library(ClustGeo)
library(rmapshaper)  # helpful functions for mapping
library(ggsn)
library(tmap)  # map viewer
library(ggmap)  # map viewer, googlemap (need to register)
library(leaflet)  # map viewer
library(mapview)  # map viewer

# Descriptive Statistics
library(naniar)
library(psych)
library(gtsummary)
```

## Environment
```{r Environment}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")

fn.prefix <- "_"
# icon.grob <- png_to_grob()

# https://en.wikipedia.org/wiki/World_Geodetic_System#WGS84
crs.base <- "+proj=longlat +datum=WGS84"
loc.base <- c(lon = 139.73, lat = 35.70)

col.bg <- "#262A33"
col.fg <- "#393F4D"
```

# Load DataSets {.tabset .tabset-fade .tabset-pills}

https://nlftp.mlit.go.jp/ksj/gml/datalist/KsjTmplt-L01-v3_1.html

Land Price Open Dataset
  
L01_006/_100: Price (JPY/m$^{2}$)  
L01_007: Price Change (%)  
L01_022: District Code  
L01_023: City  
L01_024: Address  
  
  
```{r Load DataSets}
admin.sf <- sf::read_sf(
    "input/Administrative_district_Tokyo_2021/N03-21_13_210101.shp"
) %>% 
    rmapshaper::ms_filter_islands(min_area = 1e8) %>% 
    st_transform(crs = crs.base)  # JGD2011 => WGS84

lp.sf <- sf::read_sf("input/Land-Price_Tokyo_2022/L01-22_13.shp") %>% 
    st_transform(crs = crs.base)  # JGD2011 => WGS84


admin.code <- readxl::read_excel(
  "input/AdminiBoundary_CD.xlsx", skip = 1
) %>% rename("District" = "行政区域コード")
```


# Preprocessing {.tabset .tabset-fade .tabset-pills}
## Concat Polygons by District
```{r}
admin.sf <- admin.sf %>% 
  rename(District = N03_007) %>% 
  select(District) %>% 
  group_by(District) %>% 
  summarise(.groups = "drop") %>% 
  left_join(admin.code %>% select(District, City), by = "District")
```

## Transform and Rename
```{r}
city.rm <- c(
    "Ooshima", 
    "Niijima", 
    "Kouzushima", 
    "Miyake", 
    "Hachijyou", 
    "Ogasawara"
)

lp.sf <- lp.sf %>% mutate(
    Price = L01_006 / 1e6,
    Price_Change = L01_007, 
) %>% rename(
    District = L01_022,
    District.Fine = L01_023,
    Address = L01_024
) %>% 
    left_join(admin.code %>% select(District, City), by = "District") %>% 
    filter(!City %in% city.rm)
```

# Price Statistics by Dstrict
```{r}
lp.sf.stat <- lp.sf %>% select(District, City, Price, Price_Change) %>% 
    filter(Price < 4) %>% 
    group_by(District) %>% 
    summarise(
        Price.Mean = round(mean(Price, na.rm = TRUE), 3),
        Price.Median = round(median(Price, na.rm = TRUE), 3),
        Price.Std = round(sd(Price, na.rm = TRUE), 3),
        Price_Change.Mean = round(mean(Price_Change, na.rm = TRUE), 3),
        Price_Change.Median = round(median(Price_Change, na.rm = TRUE), 3),
        Price_Change.Std = round(sd(Price_Change, na.rm = TRUE), 3),
        .groups = "drop"
    ) %>%
    st_set_geometry(NULL) %>%
    left_join(admin.sf %>% select(District), by = "District") %>%
    st_set_geometry(value = "geometry")

view.table(lp.sf.stat %>% st_set_geometry(NULL))
tbl_summary(lp.sf.stat %>% st_set_geometry(NULL) %>% select(-District))
```


# Exploratory Data Analysis {.tabset .tabset-fade .tabset-pills}
## Bad Map
```{r}
p1 <- ggplot(data = lp.sf %>% select(Price)) + 
    geom_sf(inherit.aes = FALSE, 
            mapping = aes(), 
            data = admin.sf %>% 
                filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
                select(geometry), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(aes(color = Price), size = 0.5, alpha = 0.5) + 
    scale_color_carto_c(palette = "Teal") +
    labs(x = NULL, y = NULL, 
         title = "Bad Land Price Map of Tokyo (JPY 1e6 / m^2)", 
         subtitle = "The color bar is not scaled properly, and the points do not convey good information.") +
    theme(
        title = element_text(color = col.bmc.gray.p, size = 10), 
        plot.subtitle = element_text(color = col.bmc.gray.p, size = 9), 
        axis.text = element_blank(), 
        legend.position = c(0.1, 0.20),
        panel.grid.major = element_blank(), 
        plot.background = element_rect(fill = col.bg, color = col.bg), 
        panel.background = element_rect(fill = col.bg, color = col.bg)
    ) +
    guides(color = guide_colourbar(
        title = "Price", 
        title.theme = element_text(
            color = col.bmc.gray.p, family = font.base, face = font.face, 
            size = 6
        ), title.hjust = -1, 
        barwidth = 0.4, barheight = 3, 
        label.theme = element_text(
            color = col.bmc.gray.p, family = font.base, face = font.face, 
            size = 7
        ), 
        label.position = "right", 
        ticks.linewidth = 3, ticks.colour = col.bg
    ))

p2 <- ggplot(data = lp.sf.stat) + 
    geom_sf(inherit.aes = FALSE, 
            mapping = aes(), 
            data = admin.sf %>% 
                filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
                select(geometry), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(aes(fill = Price.Mean), color = NA, size = 0.5, alpha = 0.5) + 
    scale_fill_carto_c(palette = "Teal") + 
    labs(x = NULL, y = NULL, 
         title = NULL, 
         subtitle = "In the above case, an aggregation by region can sometimes visualize to get a better idea of \nthe overall trend. If the number of data for a given region is small for aggregation, \nthere is a high uncertainty, and better stratification should be considered (e.g. spatial clustering).") +
    theme(
        title = element_text(color = col.bmc.gray.p, size = 11), 
        plot.subtitle = element_text(color = col.bmc.gray.p, size = 9), 
        axis.text = element_blank(), 
        legend.position = c(0.1, 0.175),
        panel.grid.major = element_blank(), 
        plot.background = element_rect(fill = col.bg, color = col.bg), 
        panel.background = element_rect(fill = col.bg, color = col.bg)
    ) +
    guides(fill = guide_colourbar(
        title = "Mean price\n   by region", 
        title.theme = element_text(
            color = col.bmc.gray.p, family = font.base, face = font.face, 
            size = 6
        ), title.hjust = -10, 
        barwidth = 0.4, barheight = 3, 
        label.theme = element_text(
            color = col.bmc.gray.p, family = font.base, face = font.face, 
            size = 7
        ), 
        label.position = "right", 
        ticks.linewidth = 3, ticks.colour = col.bg
    ))

p1 / p2
ggsave("fig/Tokyo-land-price.1.jpg", width = 6, height = 7)
```


## Choropleth Map: Tokyo
```{r, fig.height=15}
p1 <- lp.sf.stat %>% 
    ggplot() + 
    geom_sf(mapping = aes(), data = admin.sf, 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price.Mean), color = "white", 
                 cex = 2.5, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "Mean Price (1e6 JPY)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 30, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 10, label.position = "right", 
        ticks.linewidth = 3
    ))

p2 <- lp.sf.stat %>% 
    ggplot() + 
    geom_sf(mapping = aes(), data = admin.sf, 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Median)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price.Median), color = "white", 
                 cex = 2.5, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "Median Price (1e6 JPY)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 30, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 10, label.position = "right", 
        ticks.linewidth = 3
    ))

p3 <- lp.sf.stat %>% 
    ggplot() + 
    geom_sf(mapping = aes(), data = admin.sf, 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price_Change.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "ArmyRose") + 
    geom_sf_text(aes(label = Price_Change.Mean), color = "white", 
                 cex = 2.5, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "Mean Price Change (%)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 10, label.position = "right", 
        ticks.linewidth = 3
    ))

p1 / p2 / p3
# ggsave("fig/Land-price_Tokyo.jpg", dpi = 300, width = 10, height = 15)
```

## Choropleth Map: Central Tokyo
```{r, fig.height=24, fig.width=8}
p1 <- lp.sf.stat %>% filter(District <= 13123) %>% 
    ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf %>% filter(District <= 13123), 
          fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf_text(aes(label = City), color = col.os, 
               cex = 4, family = "Candara") + 
  labs(x = NULL, y = NULL, title = "Central Tokyo Districts") + 
  theme(title = element_text(size = 15), 
        plot.margin = unit(c(0, 0, 40, 0), units = "pt"))

p2 <- lp.sf.stat %>% filter(District <= 13123) %>% 
  ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf %>% filter(District <= 13123), 
          fill = NA, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf(mapping = aes(fill = Price.Median), color = "white") + 
  geom_sf_text(aes(label = Price.Median), color = col.os, 
               cex = 4, family = "Candara") + 
  rcartocolor::scale_fill_carto_c(name = "", palette = "Tropic") + 
  labs(x = NULL, y = NULL, title = "Median Land Price (1e6 JPY)") + 
  theme(title = element_text(size = 15), 
        legend.position = c(0.05, 0.3), 
        plot.margin = unit(c(0, 0, 40, 0), units = "pt")) + 
  guides(fill = guide_colourbar(
    barwidth = 0.5, barheight = 15, label.position = "right", 
    ticks.linewidth = 3
  ))

p3 <- lp.sf.stat %>% filter(District <= 13123) %>% 
  ggplot() + 
  geom_sf(mapping = aes(), data = admin.sf %>% filter(District <= 13123), 
          fill = NA, color = col.cb, alpha = 0.1, size = 0.1) + 
  geom_sf(mapping = aes(fill = Price_Change.Mean), color = "white") + 
  geom_sf_text(aes(label = Price_Change.Mean), color = col.os, 
               cex = 4, family = "Candara") + 
  rcartocolor::scale_fill_carto_c(name = "", palette = "ArmyRose") + 
  labs(x = NULL, y = NULL, title = "Mean Land Price Change (%)") + 
  theme(title = element_text(size = 15), 
        legend.position = c(0.05, 0.3), 
        plot.margin = unit(c(0, 0, 0, 0), units = "pt")) + 
  guides(fill = guide_colourbar(
    barwidth = 0.5, barheight = 15, label.position = "right", 
    ticks.linewidth = 3
  ))

p1 / p2 / p3
# ggsave("fig/Land-price_Central-Tokyo.jpg", dpi = 300, width = 10, height = 30)
```


# Spatial Clustering with ClustGeo
## Standardize Node Features
```{r}
lp.stat.scale <- lp.sf.stat %>% 
    as.data.frame() %>% 
    select(-District, -City, -Price_Change.Median, -geometry) %>% 
    scale() %>% as.data.frame()
```

## Compute Neighboorhod list with `spdep::poly2nb`
Works fine with `MULTIPOLYGON`, but not with NA.
```{r}
lp.stat.nb <- spdep::poly2nb(lp.sf.stat %>% select(geometry))
```

## Compute dissimilarities
```{r}
A <- spdep::nb2mat(lp.stat.nb, style = "B")
diag(A) <- 1
colnames(A) <- rownames(A) <- rownames(lp.stat.scale)
```

## Optimize alpha
```{r}
D0 <- dist(lp.stat.scale)
D1 <- as.dist(1 - A)
alpha.seq <- seq(0, 1, 0.1)
K <- 5
cr <- choicealpha(D0, D1, alpha.seq, K, 
                  wt = NULL, scale = TRUE, graph = FALSE)
cr$Q  # proportion of explained inertia
crQ.df <- data.frame(cr$Qnorm) %>% mutate(alpha = alpha.seq)
rownames(crQ.df) <- NULL
```

```{r}
crQ.df %>% ggplot() + 
  geom_point(mapping = aes(x = alpha, y = Q0norm), 
             color = col.plos.yellow, size = 2) + 
  geom_line(mapping = aes(x = alpha, y = Q0norm), 
            color = col.plos.yellow, size = 1) + 
  geom_point(mapping = aes(x = alpha, y = Q1norm), 
             color = col.plos.pink, size = 2) + 
  geom_line(mapping = aes(x = alpha, y = Q1norm), 
            color = col.plos.pink, size = 1) + 
  lims(y = c(0, NA)) + 
  labs(y = "Q: proportion of explained inertia", 
       title = "ClustGeo::choicealpha") + 
  theme(plot.title = element_text(size = 15))
```

## Regionalization {.tabset .tabset-fade .tabset-pills}
```{r}
tree <- 
  ClustGeo::hclustgeo(D0 = D0, D1 = D1, alpha = 0.2, scale = TRUE, wt = NULL)
```

### ncluster = 3
```{r}
lp.sf.stat <- lp.sf.stat %>% 
  mutate(partition.3 = factor(cutree(tree = tree, k = 3), 
                              levels = 1:3, labels = 1:3))
table(lp.sf.stat$partition.3)
```

```{r, fig.height=20}
p1 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Median)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    scalebar(
        data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
        location = "bottomright",  # waved if anchor set.
        anchor = setNames(c(139.3, 35.52), c("x", "y")), 
        dist = 10, dist_unit = "km",
        # Esri Japan: https://bit.ly/3LtcBaE
        transform = TRUE, model = "WGS84",
        height = 0.01, st.dist = 0.03, st.size = 4, st.color = col.os, 
        box.fill = c(col.tw, col.plos.pink), 
        box.color = "white", border.size = 0.05, 
        family = font.base
    ) + 
    geom_sf_text(aes(label = Price.Median), color = "white",
                 cex = 5, family = font.base) + 
    labs(x = NULL, y = NULL, title = "Median Price (1e6 JPY)") + 
    theme(title = element_text(size = 15), 
          # legend.position = c(0.05, 0.35), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p2 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price_Change.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price_Change.Mean), color = "white",
                 cex = 5, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "Mean Price Change (%)") + 
    theme(title = element_text(size = 15), 
          # legend.position = c(0.05, 0.35), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p3 <- lp.sf.stat %>% ggplot() + 
    geom_sf(mapping = aes(fill = partition.3), size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 4, family = font.base) + 
    labs(x = NULL, y = NULL, title = "ncluster = 3") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt"))

p1 / p2 / p3 + 
    plot_annotation(
    title = "Spatial Clustering on Price Stats", 
    subtitle = "", 
    theme = theme(title = element_text(size = 20, family = "Candara"))
    )

ggsave("fig/Spatial-Clustering-Tokyo_ClustGeo_n3.jpg", 
       dpi = 300, width = 15, height = 10 * 3)
```

### ncluster = 5
```{r}
lp.sf.stat <- lp.sf.stat %>% 
  mutate(partition.5 = factor(cutree(tree = tree, k = 5), 
                              levels = 1:5, labels = 1:5))
table(lp.sf.stat$partition.5)
```

```{r, fig.height=20}
p1 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price.Median)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    scalebar(
        data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
        location = "bottomright",  # waved if anchor set.
        anchor = setNames(c(139.3, 35.52), c("x", "y")), 
        dist = 10, dist_unit = "km",
        # Esri Japan https://bit.ly/3LtcBaE
        transform = TRUE, model = "WGS84",
        height = 0.01, st.dist = 0.03, st.size = 4, st.color = col.os, 
        box.fill = c(col.tw, col.plos.pink), 
        box.color = "white", border.size = 0.05, 
        family = font.base
    ) +   
    geom_sf_text(aes(label = Price.Median), color = "white",
                 cex = 5, family = "Candara") +
    labs(x = NULL, y = NULL, title = "Median Price (1e6 JPY)") + 
    theme(title = element_text(size = 15),
          # legend.position = c(0.05, 0.35), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p2 <- lp.sf.stat %>% 
    filter(!City %in% c("Oku-Tama", "Hinohara")) %>% 
    ggplot() + 
    geom_sf(mapping = aes(), 
            data = admin.sf %>% filter(!City %in% c("Oku-Tama", "Hinohara")), 
            fill = col.cb, color = col.cb, alpha = 0.1, size = 0.1) + 
    geom_sf(mapping = aes(fill = Price_Change.Mean)) + 
    rcartocolor::scale_fill_carto_c(name = "", palette = "Magenta") + 
    geom_sf_text(aes(label = Price_Change.Mean), color = "white",
                 cex = 5, family = "Candara") +
    labs(x = NULL, y = NULL, title = "Mean Price Change (%)") + 
    theme(title = element_text(size = 15), 
          legend.position = c(0.05, 0.35), 
          plot.margin = unit(c(0, 0, 20, 0), units = "pt")) + 
    guides(fill = guide_colourbar(
        barwidth = 0.5, barheight = 12, label.position = "right", 
        ticks.linewidth = 3
    ))

p3 <- lp.sf.stat %>% ggplot() + 
    geom_sf(mapping = aes(fill = partition.5), size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 4, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "ncluster = 5") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt"))

p1 / p2 / p3 + 
    plot_annotation(
    title = "Spatial Clustering on Price Stats", 
    subtitle = "", 
    theme = theme(title = element_text(size = 20, family = "Candara"))
    )

ggsave("fig/Spatial-Clustering-Tokyo_ClustGeo_n5.jpg", 
       dpi = 300, width = 15, height = 10 * 3)
```

### ncluster = 10
```{r}
lp.sf.stat <- lp.sf.stat %>% 
  mutate(partition.10 = factor(cutree(tree = tree, k = 10), 
                               levels = 1:10, labels = 1:10))
table(lp.sf.stat$partition.10)
```

```{r, fig.height=20}
p1 <- lp.sf.stat %>% ggplot() + 
    geom_sf(mapping = aes(fill = partition.3), size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    geom_sf_text(aes(label = City), color = col.os, 
                 cex = 4, family = "Candara") + 
    labs(x = NULL, y = NULL, title = "ncluster = 3") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 30, 0), units = "pt")) + 
  annotation_custom(grob = t.grob, 
                    xmin = 139.8, xmax = 139.8, 
                    ymin = 35.83, ymax = 35.83) + 
  annotate(geom = "text", x = 139.83, y = 35.83, 
           label = "@Maxwell_110", size = 3, family = "Candara", color = "gray")

p2 <- lp.sf.stat %>% ggplot() + 
    geom_sf(mapping = aes(fill = partition.5), size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    labs(x = NULL, y = NULL, title = "ncluster = 5") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 30, 0), units = "pt"))

p3 <- lp.sf.stat %>% ggplot() + 
    geom_sf(mapping = aes(fill = partition.10), size = 0.2, alpha = 0.5) + 
    scale_fill_carto_d(name = "", palette = "Vivid") + 
    labs(x = NULL, y = NULL, title = "ncluster = 10") + 
    theme(title = element_text(size = 15), 
          legend.position = "none", 
          plot.margin = unit(c(0, 0, 0, 0), units = "pt"))

p1 / p2 / p3 + 
    plot_annotation(
    title = "Spatial Clustering on Price Stats", 
    subtitle = "(ClustGeo)", 
    theme = theme(title = element_text(size = 20, family = "Candara"))
    )

ggsave("fig/Spatial-Clustering-Tokyo_ClustGeo_n3-10.jpg", 
       dpi = 300, width = 15, height = 10 * 3)
ggsave("fig/Spatial-Clustering-Tokyo_ClustGeo_n3-10_dpi150.jpg", 
       dpi = 150, width = 15, height = 10 * 3)
```
