---
title: "Non-Life"
subtitle: "Naive Statistics"
author: ""
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: show
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

# Preparations {.tabset .tabset-fade .tabset-pills}

## R Packages
```{r, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(readxl)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# GeoSpatial Analysis
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping
library(rmapshaper)  # helpful functions for mapping
library(tmap)  # map viewer
library(leaflet)  # map viewer
library(mapview)  # map viewer

# Parallel
library(purrr)
library(foreach)
library(doParallel)

# Descriptive Statistics
library(naniar)
library(psych)
library(gtsummary)

# Multi Level Model
library(lme4)
library(lmerTest)
library(coefplot)

# Stan
library(rstan)
library(bayesplot)
library(broom)
library(broom.mixed)
```

## Environment
```{r Environment}
# rgb(73, 81, 112, maxColorValue = 255): "#495170"
col.os <- "#414a4c"  # Outer Space
col.rp <- "#7851a9"  # Royal Purple
col.cb <- "#b0b7c6"  # Cadet Blue
col.el <- "#ceff1d"  # Electric Lime
col.rm <- "#e3256b"  # Razzmatazz
col.plos.ye <- "#D6E13D"  # PLOS ONE Yellow
col.plos.pink <- "#CF00A3"  # PLOS ONE Pink
scale.col.1 <- c()
font.base <- "Candara"
theme_set(theme_minimal(base_family = font.base))
options(dplyr.summarise.inform = TRUE)
# Twitter Logo and Acc
t.img <- png::readPNG("fig/twitter.png")
t.grob <- grid::rasterGrob(t.img, 
                           width = unit(0.4, "cm"), 
                           height = unit(0.4, "cm"),
                           interpolate = FALSE)
```

# Load DataSets {.tabset .tabset-fade .tabset-pills}

\singlespacing

Data sorce: TheGeneral Insurance Association of Japan  
https://www.sonpo.or.jp/report/statistics/syumoku/index.html

\doublespacing

| **Format of excel files on a fiscal year basis**:  
|    Sheet 1: Written Premium by Large Coverage  
|    Sheet 2: Written Premium by Small Coverage  
|    Sheet 3:  
|    Sheet 4:  
|    Sheet 5:  
|    Sheet 6:  
|    Sheet 7:  
| 
| **Format of an excel file of net primary premium transition**:  
|    Sheet 1: 1994 - latest year (2020)

\singlespacing

```{r Load DataSets}
base.input.path <- "input/Nonlife/"
excel.names <- sort(list.files(base.input.path))
excel.abs.paths <- paste0(base.input.path, excel.names)

net.premium.all <- readxl::read_excel(
   path = excel.abs.paths[length(excel.abs.paths)], 
   skip = 0  # w/o cleansing skip should be 2.
) %>% slice(-1)
```

# Preprocessing
## Rename Columns
```{r}
col.names <- c("Coverage")
for (yr in 1994:2020) {
   col.names <- c(col.names, paste0(yr, c("_MJPY", "_RATIO")))
}
colnames(net.premium.all) <- col.names
```

## Coverage in En
```{r}
net.premium.all <- net.premium.all %>% 
   mutate(
      Coverage_En = c(
         "Fire", "Auto", "Accident", "Misc", 
         # Marine-Transit = Hull + Marine-Cargo + Transit
         "Marine-Transit", "Hull", "Marine-Cargo", "Transit", 
         "Subtotal", "Mandatory-Auto-Liability", "Sum"
      )
   )
```

## Long Format
```{r}
net.premium.all.long <- net.premium.all %>% 
   select(Coverage_En, contains("MJPY")) %>% 
   # filter(!Coverage_En %in% 
   #           c("Hull", "Marine-Cargo", "Transit", "Subtotal")) %>% 
   pivot_longer(cols = -Coverage_En, 
                names_to = "YEAR", values_to = "NET_PREMIUM") %>% 
   mutate(
      YEAR = as.integer(str_remove(YEAR, pattern = "_MJPY")), 
      NET_PREMIUM = as.numeric(NET_PREMIUM)
   )
```


# Yearly Changes in Net Premium
```{r, fig.height = 8}
net.premium.all.long %>% 
   filter(!Coverage_En %in% 
             c("Hull", "Marine-Cargo", "Transit", "Subtotal")) %>% 
   ggplot() + 
   geom_point(aes(x = YEAR, y = NET_PREMIUM, color = Coverage_En)) + 
   geom_line(aes(x = YEAR, y = NET_PREMIUM, color = Coverage_En)) + 
   scale_color_carto_d(
      palette = "Pastel", 
      guide = guide_legend(title = expression("Coverage: "))
   ) + 
   annotation_custom(grob = t.grob, 
                     xmin = 2000.5, xmax = 2000.5, 
                     ymin = 8.1e6, ymax = 8.1e6) + 
   annotate(geom = "text", x = 2002.2, y = 8.12e6, 
            label = "@Maxwell_110", size = 3.5, family = font.base) + 
   labs(x = NULL, 
        y = "Net Premium (1e6 JPY)", 
        title = "Yearly Changes in Net Premium", 
        subtitle = "Data sorce: TheGeneral Insurance Association of Japan:\n https://www.sonpo.or.jp/report/statistics/syumoku/index.html") + 
   theme(legend.position = c(0.85, 0.675), 
         plot.title = element_text(size = 20, 
                                   margin = margin(t = 40, b = -40)), 
         plot.subtitle = element_text(size = 10, 
                                      margin = margin(t = 50, b = -50)))
ggsave("fig/non-life_net-premium-change.jpg", 
       width = 10, height = 8)
```

