---
title: "shapviz"
subtitle: "Vignettes"
author: "Maxwell"
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 15
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

https://cran.r-project.org/web/packages/shapviz/vignettes/shapviz.html

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

library(shapviz)

library(xgboost)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")
# icon.grob <- png_to_grob()

theme_update(panel.grid.minor = element_blank())
```


# Dataset {.tabset .tabset-fade .tabset-pills}
```{r Dataset}
data("penguins", package = "modeldata")  # modeldata from tidymodels
glimpse(penguins)
```

```{r}
p <- ggmice::plot_pattern(data = penguins, vrb = "all", square = TRUE, 
                          rotate = FALSE, cluster = NULL)
p + theme_minimal(base_family = font.base)
```


# Remove Examples with Missing
```{r}
penguins <- penguins %>% filter(!is.na(sex))
```


# Fit an XGBoost Model
```{r}
X <- penguins %>% select(-body_mass_g)
dtrain <- xgb.DMatrix(data.matrix(X), label = penguins$body_mass_g)

fit <- xgb.train(
  params = list(learning_rate = 0.1, objective = "reg:squarederror"), 
  data = dtrain,
  nrounds = 65L
)
```


# Create `shapviz` Object
One line of code creates a shapviz object. It contains SHAP values and feature values for the set of observations we are interested in.  
Note again that X is solely used as explanation dataset, not for calculating SHAP values.  
In this example, we construct the shapviz object directly from the fitted XGBoost model. Thus we also need to pass a corresponding prediction dataset X_pred used for calculating SHAP values by XGBoost.

```{r}
X.s <- X[sample(nrow(X), 200), ]

# X is the "explanation" dataset using the original factors
shp <- shapviz(fit, X_pred = data.matrix(X.s), X = X.s)
```


# sv_waterfall(), sv_importance(), and sv_dependence()
Waterfall plots to study single predictions.  
Options (...) for `ggfittext::geom_fit_text` can be allowed.  
```{r fig.height=20}
# ggplot class
p1 <- sv_waterfall(shp, 
                   row_id = 2, 
                   max_display = 10L,
                   sort_fun = NULL,
                   fill_colors = c(col.bmc.sky, col.bmc.pink),
                   # format_shap = getOption("shapviz.format_shap"),
                   format_shap = function(z)
                       prettyNum(z, digits = 3, scientific = FALSE),
                   format_feat = getOption("shapviz.format_feat"),
                   format_fun = NULL,
                   contrast = FALSE,
                   show_connection = TRUE,
                   show_annotation = FALSE,
                   annotation_size = 3.2,
                   family = font.base,
                   color = "white")
p1 <- p1 + labs(title = "shapviz::sv_waterfall()", 
                subtitle = "Using penguin dataset and XGBoost") + 
    geom_text(aes(x = 4250, y = 0.75, label = "@Maxwell_110"), 
              family = font.base, color = col.os, alpha = 0.05) + 
    theme_minimal(base_family = font.base) + 
    theme(text = element_text(family = font.base, color = col.os), 
          title = element_text(size = 15, color = col.os), 
          plot.title = element_text(hjust = 0.))


p2 <- sv_importance(object = shp, 
                    kind = "both", 
                    fill = col.bmc.green.l, 
                    alpha = 0.5, 
                    size = 1, 
                    bar_width = 1 / 3, 
                    viridis_args = NULL)
p2 <- p2 + 
    labs(y = "Mean SHAP value", 
         title = "shapviz::sv_importance()", 
         subtitle = "kind = both") + 
    # A: magma, B: inferno, C: plasma, D: viridis, E: cividis
    scale_color_viridis_c(option = "E") + 
    theme_minimal(base_family = font.base) + 
    theme(text = element_text(family = font.base, color = col.os), 
          title = element_text(size = 15, color = col.os), 
          plot.title = element_text(vjust = 0.), 
          legend.title = element_blank())

p3 <- sv_dependence(object = shp, 
                    v = "flipper_length_mm", 
                    color_var = "species", 
                    size = 3, alpha = 0.5)
p3 <- p3 + 
    labs(title = "shapviz::sv_dependence()") + 
    scale_fill_carto_d(palette = "Safe") + 
    theme_minimal(base_family = font.base) + 
    guides(colour = guide_legend(
        title = "Penguin species", 
        override.aes = list(size = 7)
    )) + 
    theme(text = element_text(family = font.base, color = col.os), 
          title = element_text(size = 15, color = col.os), 
          plot.title = element_text(vjust = 0.))

p1 + plot_spacer() + p2 + p3 + plot_spacer() + 
    plot_layout(ncol = 3, widths = c(6, 0.5, 5), heights = c(2, 1.5))
ggsave("fig/shapviz-01.jpg", width = 20, height = 15)
```


# References


