---
title: "Generalized Linear Mixed Models"
subtitle: "Based on C. Manning, 2007 paper"
author: "HoxoMaxwell"
date: '`r Sys.Date()`'
output: 
   html_document:
      number_sections: true
      toc: true
      toc_depth: 5
      toc_float: true
      theme: cosmo
      highlight: tango
      code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

This notebook is the followup study of  
https://nlp.stanford.edu/manning/courses/ling289/GLMM.pdf  

# Load Library
```{r}
library(languageR)
library(DT)
library(summarytools)
library(tidyverse)
library(lme4)
library(lmerTest)
```

# Load Data
```{r}
data("dative")
datatable(
  dative,
  options = list(pageLength = 10, 
                 autoWidth = TRUE,
                 scrollX = TRUE,
                 columnDefs = list(
                   list(width = '75px',
                        targets = seq(1, dim(dative)[2])))),
  class = 'cell-border stripe'
  )
```

- Speaker: factor  
coding speaker; available only for the subset of spoken English.

- Modality: factor  
with levels spoken, written.

- Verb: factor  
with the verbs as levels.

- SemanticClass: factor  
with levels a (abstract: 'give it some thought'), c (communication: 'tell, give me your name'), f (future transfer of possession: 'owe, promise'), p (prevention of possession: 'cost, deny'), and t (transfer of possession: 'give an armband, send').

- LengthOfRecipient: numeric  
vector coding the number of words comprising the recipient.

- AnimacyOfRec: factor  
with levels animate and inanimate for the animacy of the recipient.

- DefinOfRec: factor  
with levels definite and indefinite coding the definiteness of the recipient.

- PronomOfRec: factor  
with levels nonpronominal and pronominal coding the pronominality of the recipient.

- LengthOfTheme: numeric  
vector coding the number of words comprising the theme.

- AnimacyOfTheme: factor  
with levels animate and inanimate coding the animacy of the theme.

- DefinOfTheme: factor  
with levels definite and indefinite coding the definiteness of the theme.

- PronomOfTheme: factor  
with levels nonpronominal and pronominal coding the pronominality of the theme.

- RealizationOfRecipient: factor  
with levels NP and PP coding the realization of the dative.

- AccessOfRec: factor  
with levels accessible, given, and new coding the accessibility of the recipient.

- AccessOfTheme: factor  
with levels accessible, given, and new coding the accessibility of the theme.

# EDA
```{r, results='asis'}
dfSummary(dative, plain.ascii = F, style = "grid", graph.magnif = .75)
```

## Extract Subset
```{r}
spdative <- dative %>% filter(Modality == "spoken")
```

## xtabs and mosaicplot
```{r}
spdative.xtabls <- xtabs(
    formula =  ~ RealizationOfRecipient + AnimacyOfTheme + AnimacyOfRec,
    data = spdative
    )
spdative.xtabls
```
```{r}
mosaicplot(spdative.xtabls, 
           main = "mosaicplot",
           color="slategray")
```

# FE
```{r}
spdative <- spdative %>%
    mutate(LengthOfThemeMinusRecipient = LengthOfTheme - LengthOfRecipient) %>%
    mutate(RatioOfLengthsThemeOverRecipient = LengthOfTheme / LengthOfRecipient)
```

```{r}
divs1 <- c(0, 1, 2, 3, 4, 5, 6, 10, 100)
divs2 <- c(-40, -6, -4, -3, -2, -1, 0, 1, 2, 3, 4, 6, 40)
divs3 <- c(-2.5, -1.5, -1, -0.5, -0.25, 0, 0.5, 1, 1.5, 2, 2.5, 4)
recip.leng.group <- cut(spdative$LengthOfRecipient, divs1)
theme.leng.group <- cut(spdative$LengthOfTheme, divs1)
theme.minus.recip.group <- cut(spdative$LengthOfThemeMinusRecipient, divs2)
log.theme.over.recip.group <-
    cut(log(spdative$RatioOfLengthsThemeOverRecipient), divs3)
recip.leng.table <- table(recip.leng.group, 
                          spdative$RealizationOfRecipient)
recip.leng.table
```

# GLMM
## Fit1
```{r}
dative.glmm1 <- glmer(
    formula = RealizationOfRecipient ~ 
        SemanticClass + LengthOfRecipient + AnimacyOfRec + DefinOfRec+ 
        PronomOfRec + AccessOfRec + LengthOfTheme + AnimacyOfTheme + 
        DefinOfTheme + PronomOfTheme + AccessOfTheme +
        (1 | Speaker) + (1 | Verb),
    family = "binomial",
    data = spdative
    )
summary(dative.glmm1)
```

## Fit2
- remove a random effect on Speaker
```{r}
dative.glmm2 <- glmer(
    formula = RealizationOfRecipient ~ 
        SemanticClass + LengthOfRecipient + AnimacyOfRec + DefinOfRec+ 
        PronomOfRec + AccessOfRec + LengthOfTheme + AnimacyOfTheme + 
        DefinOfTheme + PronomOfTheme + AccessOfTheme +
        (1 | Verb),
    family = "binomial",
    data = spdative
    )
summary(dative.glmm2)
```

## Fit3
```{r}
dative.glmm3 <- glmer(
    formula = RealizationOfRecipient ~ 
        SemanticClass + LengthOfRecipient + AnimacyOfRec + DefinOfRec+ 
        PronomOfRec + AccessOfRec + LengthOfTheme + AnimacyOfTheme + 
        DefinOfTheme + PronomOfTheme + AccessOfTheme +
        log(RatioOfLengthsThemeOverRecipient) +
        (1 | Verb),
    family = "binomial",
    data = spdative
    )
summary(dative.glmm3)
```


# EOF
