---
title: "disbayes"
subtitle: "Vignette 01"
author: "Maxwell"
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 15
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

https://cran.r-project.org/web/packages/disbayes/vignettes/disbayes.html

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(ggsci)  # Journal palette
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)

library(disbayes)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")

fn.prefix <- "disbayes-01_"
# icon.grob <- png_to_grob()
```


# Dataset {.tabset .tabset-fade .tabset-pills}
`ihdengland`  
Ischemic heart disease in England  
sex: "male" or "female".  
ageyr: Year of age.  
location: Name of the location, which is either a city region or region in England.  
num_mort: Numerator behind the estimate of mortality  
num_inc: Numerator behind the estimate of incidence  
num_prev: Numerator behind the estimate of prevalence  
denom_mort: Denominator behind the estimate of mortality  
denom_inc: Denominator behind the estimate of incidence  
denom_prev: Denominator behind the estimate of prevalence  

```{r Dataset}
data("ihdengland", package = "disbayes")
glimpse(ihdengland)
gtsummary::tbl_summary(ihdengland, by = gender)
```

```{r}
ihdbristol <- ihdengland %>% filter(area=="Bristol", gender=="Male")
ihdbristol %>% filter(between(age, 50, 55))
```


# Running disbayes
eqage: Case fatalities (and incidence and remission rates) are assumed to be equal for all ages below this age, inclusive, when using the smoothed model.  

```{r}
dbres <- disbayes(
    data = ihdbristol, 
    age = "age", 
    inc_num = "inc_num", inc_denom = "inc_denom", 
    prev_num = "prev_num", prev_denom = "prev_denom",
    mort_num = "mort_num", mort_denom = "mort_denom", 
    eqage = 40
)
```

```{r}
dbres.tidy <- tidy(dbres) %>% 
    rename(
        p2.5 = `2.5%`, 
        p25 = `25%`, 
        p50 = `50%`,
        p75 = `75%`, 
        p97.5 = `97.5%`
    )
glimpse(dbres.tidy)

p <- ggmice::plot_pattern(data = dbres.tidy, square = TRUE)
p + theme_minimal(base_family = font.base)
```


# Viz
```{r}
# plot(dbres)
dbres.tidy %>% filter(var == "cf") %>% 
    ggplot() + 
    geom_pointrange(aes(x = age, y = p50, ymin = p2.5, ymax = p97.5), 
                    size = 0.1, color = col.plos.pink) + 
    labs(y = NULL, title = "Estimated case fatality for IHD in England:", 
         subtitle = "using R::disbayes package", 
         caption = "@ Maxwell_110") + 
    theme(
        title = element_text(size = 10), 
        plot.subtitle = element_text(size = 10), 
        plot.caption = element_text(size = 7, color = "gray77"), 
        axis.title = element_text(size = 10)
    )
ggsave(paste0("fig/", fn.prefix, "case-fatality.jpg"), 
       width = 6, height = 4, dpi = 300)
```


# References
https://chjackson.github.io/disbayes/articles/disbayes.html
