---
title: "Bayesian structure and Causal learning of Gaussian DAGs"
subtitle: "R package: BCDAG vignettes vol.1"
author: ""
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
abstract: \singlespacing This is the first of a series of three vignettes introducing the R package BCDAG. In this vignette we focus on functions `rDAG()` and `rDAGWishart()` which implement random generation of DAG structures and DAG parameters under the assumption that the joint distribution of variables X_1, ..., X_q is Gaussian and the corresponding model (Choleski) parameters follow a DAG-Wishart distribution. Finally, data generation from Gaussian DAG models is described.
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: show
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

<br>

<br>

**Random data generation from Gaussian DAG models**  
https://cran.r-project.org/web/packages/BCDAG/vignettes/bcdag_generatedata.html

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)

# Parallel
library(purrr)
library(foreach)
library(doParallel)

# BCDAG
library(ggraph)
library(tidygraph)
library(BCDAG)

source(file = "utility/helper_functions.R")
```

## Environment
```{r Environment}
# rgb(73, 81, 112, maxColorValue = 255)
col.tw <- "#dbd7d2"  # Timberwolf
col.os <- "#414a4c"  # Outer Space
col.rp <- "#7851a9"  # Royal Purple
col.pb <- "#1ca9c9"  # Pacific Blue
col.cb <- "#b0b7c6"  # Cadet Blue
col.el <- "#ceff1d"  # Electric Lime
col.cg <- "#00cc99"  # Caribbean Green
col.mt <- "#ff8243"  # Mango Tango
col.rm <- "#e3256b"  # Razzmatazz
col.sl <- "#fc2847"  # Scarlet
col.plos.yellow <- "#D6E13D"  # PLOS ONE Yellow
col.plos.pink <- "#CF00A3"  # PLOS ONE Pink
col.bmc.gray <- "#FFFDF6"  # BMC Gray
col.bmc.purple <- "#BF12F8"  # BMC Purple
col.bmc.navy <- "#1A2E4F"  # BMC Navy Blue
col.bmc.blue <- "#014A81"  # BMC Blue
col.bmc.sky <- "#0F99BE"  # BMC Sky Blue
col.bmc.green <- "#004940"  # BMC Deep Green
col.bmc.pink <- "#EF255F"  # BMC Pink

font.base <- "Candara"
# font.base <- "Times New Roman"
theme_set(theme_minimal(base_family = font.base))
options(dplyr.summarise.inform = TRUE)
# Twitter Logo and Acc
# Usage: add a code line below
# annotation_custom(grob = icon.grob, xmin = 1, xmax = 1, ymin = 1, ymax = 1)
icon.arr <- png::readPNG("fig/twitter.png")
icon.grob <- grid::rasterGrob(icon.arr, 
                              width = unit(1.20 * 3, "cm"), 
                              height = unit(0.99 * 3, "cm"), 
                              interpolate = FALSE)

set.seed(2022)
```

# Generating a binary adjacency matrix w/ `rDAG` function
<br>
`> rDAG(q, w)`  
rDAG() can be used to randomly generate a DAG structure $D = (V, E)$, 
where $V = \{1, ..., q\}$ and $E \subseteq V \times V$.  

`Arguments`  
$q$: number of nodes  
$w$: the degree of sparsity  

`Details`  
An adjacency matrix of the generated DAG, with element 1 at position $(u, v)$ indicating the presence of an edge $u \to v$.

```{r}
q <- 20
DAG <- rDAG(q = q, w = 0.5)
class(DAG)
DAG
```

# Visualization of generated DAG via `tidygraph` and `ggraph`
Reference: https://ggplot2-book.org/networks.html#networks  

## Convert adjacency matrix to tidygraph object
A tbl_graph object has Node and Edge Data.  
```{r}
dag.tg <- tidygraph::as_tbl_graph(DAG, directed = TRUE)
class(dag.tg)
dag.tg
```

## Draw Networks
`ggraph` makes ggplot object from tbl_graph object.  
Layout **circlepack** is cool!  

```{r fig.height=20}
set_graph_style(plot_margin = margin(1, 1, 1, 1))
class(ggraph(dag.tg))

p1 <- ggraph(dag.tg, layout = "auto") + 
    geom_edge_link(
        arrow = grid::arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(3, "pt"), 
        end_cap = circle(3, "pt"), 
        color = col.os, 
        edge_width = 0.5, edge_alpha = 0.5
    ) + 
    geom_node_point(color = col.plos.pink, size = 2, alpha = 0.9) + 
    theme_minimal(base_family = font.base) + 
    labs(title = "auto", x = NULL, y = NULL) + 
    theme(
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
    )

p2 <- ggraph(dag.tg, layout = "circlepack") + 
    geom_edge_link(
        arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(3, "pt"), 
        end_cap = circle(3, "pt"), 
        color = col.os, 
        edge_width = 0.5, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.plos.pink, size = 2, alpha = 0.9) + 
    theme_minimal(base_family = font.base) + 
    labs(title = "circlepack", x = NULL, y = NULL) + 
    theme(
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
    )

p1 / p2
ggsave(filename = "fig/BCDAG-01-1.jpg", dpi = 200, width = 5, height = 10)
```


# Generating Gaussian DAG parameters w/ `rDAGWishart` function
<br>
Consider **a Gaussian DAG model** of the form  
$f(X_1, ..., X_q | L,D) \sim N_q(0, (L D^{-1} L^T)^{-1})$,  
where $(L, D)$ are model parameters providing the decomposition of the **precision (inverse-covariance) matrix** $\Omega = L D^{-1} L^T$; 
specifically, $L$ is a $(q, q)$ matrix of coefficients such that for each $(u, v)$-element $L_{uv}$ with $u \neq v$, 
we have $L_{uv} \neq 0$ if and only if $(u, v) \subset E$, 
while $L_{uu} = 1$ for each $u = 1, ..., q$; 
also, $D$ is a $(q, q)$ diagonal matrix with $(u, u)$-element $D_{uu}$. 
The latter decomposition follows from the equivalent Structural Equation Model (SEM) representation of a Gaussian DAG model:

$L^{⊤} x = \epsilon , \epsilon \sim N_q(0, D)$,

where $x = (X_1, ..., X_q)^⊤$; see also Castelletti & Mascaro (2021). This is just a  Cholesky decomposition.

Function `rDAGWishart` implements random sampling from 
$(L, D)|D \sim \mathrm{DAGWishart} (a^D_c, U)$, 
where $U$ is the `rate parameter` (a $(q, q)$ s.p.d. matrix) and $a^D_c$ (a $(q, 1)$ vector) is the `shape parameter` of the DAG-Wishart distribution. 
This class of distributions was introduced by Ben David et al. (2015) as a `conjugate prior` for Gaussian DAG model-parameters. 
In its compatible version (Peluso & Consonni, 2020), elements of the vector parameter $a^D_c$ are uniquely determined from a single common shape parameter $a \gt q - 1$.

Inputs of rDAGWishart are:  
the number of samples $n$, the underlying DAG $D$, 
the common shape parameter $a$ and the rate parameter $U$. 
Given the DAG $D$ generated before, the following example implements a single ($n = 1$) draw from a compatible DAG-Wishart distribution with parameters $a = q$, $U = I_q$:

```{r}
a <- q
U <- diag(1, q)
outDL <- rDAGWishart(n = 2, DAG, a, U)
class(outDL)
```

The output of rDAGWishart() consists of two elements, $L$ and $D$ referred above equation for Gaussian DAG.

```{r}
L <- outDL$L
class(L)
dim(L)
class(L[, , 1])
diag(L[, , 1])
```
```{r}
D <- outDL$D
class(D)
dim(D)
class(D[, , 1])
sum(D[, , 1] > 0)
```

# Generating Data from a Gaussian DAG
`> solve(a, h)`  
If missing, b is taken to be an identity matrix and solve will return the inverse of a!

```{r}
# Precision matrix: L D^-1 L^T
Omega <- L[, , 1] %*% solve(D[, , 1]) %*% t(L[, , 1])

# Covariance matrix
Sigma <- solve(Omega)
```
```{r}
# MultiVariate Normal Distribution
x.mvn <- mvtnorm::rmvnorm(n = 10, sigma = Sigma)
dim(x.mvn)
```
