---
title: "Elaborate on the output of learn_DAG() using get_ functions"
subtitle: "R package: BCDAG vignettes vol.3"
author: ""
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
abstract: \singlespacing This is the third vignettes for the R package BCDAG. In this vignette, we show `how to use the output of learn_DAG() for posterior inference on DAGs, DAG parameters, and causal effect estimation`. Specifically, we introduce the functions of the `get_ family`. Remember that the output of learn_DAG() consists of an MCMC sample from the marginal posterior distribution of DAG structures (collapse = TRUE) and the joint posterior of DAGs and DAG parameters (collapse = FALSE)
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: show
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

<br>

<br>

**Elaborate on the output of learn_DAG() using get_ functions**  
https://cran.r-project.org/web/packages/BCDAG/vignettes/bcdag_getfamily.html

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)

# Parallel
library(purrr)
library(foreach)
library(doParallel)

# BCDAG
library(ggraph)
library(tidygraph)
library(BCDAG)

library(tictoc)
source(file = "utility/helper_functions.R")
```

## Environment
```{r Environment}
# rgb(73, 81, 112, maxColorValue = 255)
col.tw <- "#dbd7d2"  # Timberwolf
col.os <- "#414a4c"  # Outer Space
col.rp <- "#7851a9"  # Royal Purple
col.mb <- "#1a4876"  # Midnight Blue
col.pb <- "#1ca9c9"  # Pacific Blue
col.cb <- "#b0b7c6"  # Cadet Blue
col.el <- "#ceff1d"  # Electric Lime
col.cg <- "#00cc99"  # Caribbean Green
col.mt <- "#ff8243"  # Mango Tango
col.rm <- "#e3256b"  # Razzmatazz
col.sl <- "#fc2847"  # Scarlet

col.bg <- "#524A4E"  # Background
col.edge <- "#FDEFF4"  # Pale Pink
col.node <- "#4ECCA3"  # Teal

col.plos.yellow <- "#D6E13D"  # PLOS ONE Yellow
col.plos.pink <- "#CF00A3"  # PLOS ONE Pink
col.bmc.gray <- "#FFFDF6"  # BMC Gray
col.bmc.purple <- "#BF12F8"  # BMC Purple
col.bmc.navy <- "#1A2E4F"  # BMC Navy Blue
col.bmc.blue <- "#014A81"  # BMC Blue
col.bmc.sky <- "#0F99BE"  # BMC Sky Blue
col.bmc.green <- "#004940"  # BMC Deep Green
col.bmc.pink <- "#EF255F"  # BMC Pink

font.base <- "Candara"
# font.base <- "Times New Roman"
theme_set(theme_minimal(base_family = font.base))
options(dplyr.summarise.inform = TRUE)
# Twitter Logo and Acc
# Usage: add a code line below
# annotation_custom(grob = icon.grob, xmin = 1, xmax = 1, ymin = 1, ymax = 1)
icon.arr <- png::readPNG("fig/twitter_green.png")
icon.grob <- grid::rasterGrob(icon.arr, 
                              width = unit(.55 * 0.5, "cm"), 
                              height = unit(0.44 * 0.5, "cm"), 
                              interpolate = FALSE)
```

# Generating Data
## Generating a binary adjacency matrix w/ `rDAG` function
```{r}
set.seed(2024)  # for RMarkDown

q <- 10
w <- 0.2
DAG <- rDAG(q = q, w = w)
class(DAG)
DAG
DAG.edge.n <- sum(DAG)
```


## Visualization of generated DAG via `tidygraph` and `ggraph`
Reference:  
- https://ggplot2-book.org/networks.html#networks  
- https://cran.r-project.org/web/packages/ggraph/vignettes/tidygraph.html


```{r fig.height=10}
dag.tg <- tidygraph::as_tbl_graph(DAG, directed = TRUE)
class(dag.tg)
dag.tg

set_graph_style(plot_margin = margin(1, 1, 1, 1))
class(ggraph(dag.tg))

p1 <- ggraph(dag.tg, layout = "fr") + 
    geom_edge_link(
        arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(8, "pt"), 
        end_cap = circle(8, "pt"), 
        color = col.edge, 
        edge_width = 0.25, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.node, size = 6, alpha = 0.9) + 
    # https://www.rdocumentation.org/packages/ggraph/versions/2.0.5/topics/geom_node_text
    geom_node_text(aes(label = name), 
                   size = 3, color = col.bg, family = font.base) +
    theme_minimal(base_family = font.base) + 
    labs(title = "Gaussian DAG Structure for Synthesis", 
         subtitle = paste0("Nodes: ", as.character(q), 
                           ", Edges: ", as.character(DAG.edge.n), 
                           " (Theoretical Edge Inclusion Ratio: 0.2)"), 
         x = NULL, y = NULL) + 
    theme(
        title = element_text(color = col.tw, size = 10), 
        plot.subtitle = element_text(color = col.tw, size = 6), 
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.background = element_rect(fill = col.bg), 
    )

p1
ggsave("fig/BCDAG-03-1.png", dpi = 300, width = 3.2, height = 3.6)
```

# Generating Synthetic Data from pre-defined DAG
## Generating Gaussian DAG parameters w/ `rDAGWishart` function
```{r}
set.seed(2024)  # for RMarkDown

a <- q
U <- diag(1, q)
outDL <- rDAGWishart(n = 2, DAG, a, U)
class(outDL)

L <- outDL$L
class(L)
dim(L)
class(L[, , 1])
diag(L[, , 1])

D <- outDL$D
class(D)
dim(D)
class(D[, , 1])
sum(D[, , 1] > 0)
```

## Generating Data from a Gaussian DAG
```{r}
# Precision matrix: L D^-1 L^T
Omega <- L[, , 1] %*% solve(D[, , 1]) %*% t(L[, , 1])

# Covariance matrix
Sigma <- solve(Omega)

# MultiVariate Normal Distribution
x.mvn <- mvtnorm::rmvnorm(n = 10, sigma = Sigma)
dim(x.mvn)
```


# MCMC Sampler for DAG Parameters
## learn_DAG
```{r}
tic()
out <- learn_DAG(
    S = 5000, 
    burn = 1000, 
    data = x.mvn, 
    # In this tutorial, these hyperparameter are known.
    a = a, U = U, w = w, 
    fast = FALSE, save.memory = FALSE, 
    # DAG structures (collapse = TRUE)
    # The joint posterior of DAGs and DAG parameters (collapse = FALSE)
    collapse = FALSE, 
    verbose = FALSE
)
toc()
```

## Outputs
```{r}
class(out)
names(out)
glimpse(out)
```

## Visualization of MCMC Sampling
```{r, fig.height=4, fig.width=8}
dim(out$Graphs)
theme_set(theme_minimal(base_family = font.base))

data.frame(
    sample.no = 1:5000, 
    inclusion.ratio = colSums(out$Graphs, dims = 2) / (q * (q - 1) / 2 + q)
) %>% ggplot() + 
    geom_point(aes(x = sample.no, y = inclusion.ratio), 
               color = col.node, size = 0.5) + 
    lims(y = c(0, 1)) + 
    labs(title = "MCMC Sampling for Edge Inclusion", 
         subtitle = paste0("Theoretical Edge Inclusion Ratio: ", 
                           as.character(w)), 
         x = "MCMC Iteration", y = "Edge Inclusion Ratio") + 
    theme(
        title = element_text(color = col.node, size = 15), 
        axis.title = element_text(color = col.node, size = 12), 
        axis.text = element_text(color = col.node, size = 10), 
        panel.grid.minor = element_line(color = col.cb, linetype = 3,
                                        size = 0.1),
        panel.grid.major = element_line(color = col.edge, linetype = 3,
                                        size = 0.1),
        plot.background = element_rect(fill = col.bg)
    ) + 
    annotation_custom(grob = icon.grob, 
                      xmin = 4100, xmax = 4100, 
                      ymin = 0.9, ymax = 0.9) + 
    annotate(geom = "text", x = 4550, y = 0.9, 
             label = "@Maxwell_110",  alpha = 0.3, 
             size = 4, family = "Candara", color = col.node)
ggsave("fig/BCDAG-03-2.png", dpi = 300, width = 8, height = 4)
```

# MCMC Diagnostics of Convergence
Before using the MCMC output for posterior inference, it is common practice to perform some convergence checks. 
Function `get_diagnostics()` provides graphical diagnostics of convergence for the MCMC output of `learn_DAG()`. 
These are based on:  
- the number of edges in the DAGs  
- the posterior probability of edge inclusion for each possible edge u -> v  
both monitored across MCMC iterations.  

Input of the function is an object of class bcdag and the output consists of:  
- a traceplot and running-mean plot of the number of edges in the DAGs (graph size)  
- a collection of traceplots of the posterior probabilities of edge inclusion computed across MCMC iterations

For each pair of distinct nodes (u, v), its posterior probability of inclusion at time s (s = 1, ..., S) is estimated as the proportion of DAGs visited by the MCMC up to time s which contain the directed edge u -> v. 
Output is organized in q plots (one for each node v = 1, ..., q), each summarizing the posterior probabilities of edges u -> v (u = 1, ..., q).

```{r}
get_diagnostics(out)
```

# Posterior inference: DAG structure learning
## summary
We now show how to perform posterior inference of DAGs from the MCMC output.  
To summarize the output of learn_DAG(), a `summary()` method for objects of class bcdag is available. When `summary()` is applied to a bcdag object, a printed message appears. This summarizes the type of bcdag object (see details on bcdag object types provided in the previous vignette) and the input arguments of learn_DAG() that generated the output. In addition, the function returns some graphical outputs representing the Median Probability DAG Model estimate (`MPM`: explained later), the estimated posterior probabilities of edge inclusion and the posterior distribution of the graph size.

```{r}
summary(out)
```

## get_edgeprobs
Function `get_edgeprobs()` computes and returns the collection of posterior probabilities of edge inclusion, arranged as a (q, q) matrix, with (u, v)-element referring to edge u -> v
```{r}
out.edgeprobs <- get_edgeprobs(out)
```

```{r}
DAG2 <- DAG
rownames(DAG) <- paste0("u", 1:nrow(DAG))
colnames(DAG) <- paste0("v", 1:ncol(DAG))
DAG.df <- reshape2::melt(DAG)

rownames(out.edgeprobs) <- paste0("u", 1:nrow(out.edgeprobs))
colnames(out.edgeprobs) <- paste0("v", 1:ncol(out.edgeprobs))
out.edgeprobs.df <- reshape2::melt(out.edgeprobs)

L1 <- abs(L[, , 1])
rownames(L1) <- paste0("u", 1:nrow(L1))
colnames(L1) <- paste0("v", 1:ncol(L1))
l1.df <- reshape2::melt(L1)
```

```{r fig.width=5, fig.height=10}
theme_set(theme_minimal(base_family = font.base))

p1 <- ggplot(data = DAG.df) + 
    geom_tile(aes(x = Var2, y = Var1, fill = value)) + 
    coord_equal() + 
    labs(x = NULL, y = NULL, title = "Edge Probability Map", 
         subtitle = expression(paste("Theoretical: ", {u %->% v}))) + 
    theme(
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(), 
        axis.text.x = element_text(size = 4, face = "italic"), 
        axis.text.y = element_text(size = 4, face = "italic"), 
        legend.position = c(0.9, 0.35)
    ) + 
    scale_fill_gradient(low = col.bg, high = col.node) + 
    guides(
        fill = guide_colourbar(
            title = element_blank(), 
            barwidth = 0.3, barheight = 5, 
            label.theme = element_text(
                family = font.base, size = 4, color = col.edge
            ), 
            label.position = "right", 
            ticks.linewidth = 2
        )
    )

p2 <- ggplot(data = out.edgeprobs.df) + 
    geom_tile(aes(x = Var2, y = Var1, fill = value)) + 
    coord_equal() + 
    labs(x = NULL, y = NULL, title = NULL, 
         subtitle = "Predicted") + 
    theme(
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(), 
        axis.text.x = element_text(size = 4, face = "italic"), 
        axis.text.y = element_text(size = 4, face = "italic"), 
        legend.position = "none"
    ) + 
    scale_fill_gradient(low = col.bg, high = col.node)


p3 <- ggplot(data = l1.df) + 
    geom_tile(aes(x = Var2, y = Var1, fill = value)) + 
    coord_equal() + 
    labs(x = NULL, y = NULL, title = "Triangular Matrix L", 
         subtitle = "(in the Cholesky Decomposition of the Precision Matrix of a Bayesian network)") + 
    theme(
        plot.subtitle = element_text(size = 5), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(), 
        axis.text.x = element_text(size = 4, face = "italic"), 
        axis.text.y = element_text(size = 4, face = "italic"), 
        legend.position = "none"
    ) + 
    scale_fill_gradient(low = col.bg, high = col.node)

p1 / p2 / p3
ggsave("fig/BCDAG-03-3.png", dpi = 300, width = 3.6, height = 3 * 3)
```


## get_MPMdag
The MPM model returned in the output of summary() can be used as `a single DAG-model estimate` and is obtained by including all edges whose posterior median probability exceeds the threshold **0.5**. The MPM can be also drew by summary() function as DAG format.  
Function `get_MPMdag()` applies to an object of class bcdag and returns the (q, q) adjacency matrix of the MPM:
```{r}
MPMdag <- get_MPMdag(out)
sum(MPMdag)
```

```{r fig.height=10}
mpm.dag.tg <- tidygraph::as_tbl_graph(MPMdag, directed = TRUE)
class(mpm.dag.tg)
mpm.dag.tg

set_graph_style(plot_margin = margin(1, 1, 1, 1))
class(ggraph(mpm.dag.tg))

p.mpm <- ggraph(mpm.dag.tg, layout = "fr") + 
    geom_edge_link(
        arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(8, "pt"), 
        end_cap = circle(8, "pt"), 
        color = col.edge, 
        edge_width = 0.25, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.node, size = 6, alpha = 0.9) + 
    geom_node_text(aes(label = name), 
                   size = 3, color = col.bg, family = font.base) +
    theme_minimal(base_family = font.base) + 
    labs(title = "Median Probability DAG Model (MPM)", 
         subtitle = paste0("Nodes: ", as.character(q), 
                           ", Edges: ", as.character(sum(MPMdag)), 
                           " (Theoretical: ", as.character(DAG.edge.n), ")"), 
         x = NULL, y = NULL) + 
    theme(
        title = element_text(color = col.tw, size = 10), 
        plot.subtitle = element_text(color = col.tw, size = 6), 
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.background = element_rect(fill = col.bg), 
    )

p.mpm
```


## get_MAPdag
As an alternative, the Maximum A Posterior DAG estimate (MAP) can be considered. This corresponds to the DAG with the highest MCMC frequency of visits and can be recovered through the function get_MAPdag():
```{r}
MAPdag <- get_MAPdag(out)
sum(MAPdag)
```

```{r fig.height=10}
map.dag.tg <- tidygraph::as_tbl_graph(MAPdag, directed = TRUE)
class(map.dag.tg)
map.dag.tg

set_graph_style(plot_margin = margin(1, 1, 1, 1))
class(ggraph(map.dag.tg))

p.map <- ggraph(map.dag.tg, layout = "fr") + 
    geom_edge_link(
        arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(8, "pt"), 
        end_cap = circle(8, "pt"), 
        color = col.edge, 
        edge_width = 0.25, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.node, size = 6, alpha = 0.9) + 
    geom_node_text(aes(label = name), 
                   size = 3, color = col.bg, family = font.base) +
    theme_minimal(base_family = font.base) + 
    labs(title = "Maximum A Posterior DAG Model (MAP)", 
         subtitle = paste0("Nodes: ", as.character(q), 
                           ", Edges: ", as.character(sum(MAPdag)), 
                           " (Theoretical: ", as.character(DAG.edge.n), ")"), 
         x = NULL, y = NULL) + 
    theme(
        title = element_text(color = col.tw, size = 10), 
        plot.subtitle = element_text(color = col.tw, size = 6), 
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.background = element_rect(fill = col.bg), 
    )

p.mpm + p.map
ggsave("fig/BCDAG-03-4.png", dpi = 300, width = 3.2 * 2, height = 3.6)
```


# Posterior inference: causal effect estimation
In this last section, we introduce functions `causaleffect()` and `get_causaleffect()`, which allow to compute and estimate causal effects between variables. Specifically, we consider the causal effect on a response variable of interest consequent to a joint intervention on a given set of variables; see also Nandy et al. (2017) and Castelletti & Mascaro (2021) for formal definitions.

For a given DAG, it is possible to identify and estimate the causal effect on a node Y consequent to a hypothetical hard intervention on node j using the rules of the **do-calculus** (Pearl, 2000).  
A simple implementation of this set of rules and an estimation method for the causally sufficient case and for Gaussian data is provided by function `causaleffect()`.  
The function takes as input;  
- a numerical vector representing the labels of the intervened nodes (also called intervention target)  
- a numerical value indicating the response variable  
- the DAG model parameters L and D  
see also Castelletti & Mascaro (2021) or our previous vignette for a detailed model description.

## Understand the Difference between Single and Joint Interventions
To better understand the difference between single and joint interventions, consider as an example the total causal effect on node Y = 1 (i.e. variable X_1) of a joint intervention on nodes X_i and X_j. And the total causal effects of two separate interventions on nodes X_i and X_j under the causal model represented by the DAG generated before:
```{r}
# Joint
causaleffect(targets = c(2, 7), response = 1, 
             L = L[, , 1], D = D[, , 1])
# Single
causaleffect(targets = c(2), response = 1, 
             L = L[, , 1], D = D[, , 1])
causaleffect(targets = c(7), response = 1, 
             L = L[, , 1], D = D[, , 1])
```

The total causal effects produced are thus exactly the same for both variables in the two cases. However, if we slightly modify the DAG by adding an edge from node i to node j, so that:

```{r fig.height=10}
# i = 7, j = 2
DAG2[7, 2] <- 1
dag2.tg <- tidygraph::as_tbl_graph(DAG2, directed = TRUE)

set_graph_style(plot_margin = margin(1, 1, 1, 1))

ggraph(dag2.tg, layout = "fr") + 
    geom_edge_link(
        arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(8, "pt"), 
        end_cap = circle(8, "pt"), 
        color = col.edge, 
        edge_width = 0.25, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.node, size = 6, alpha = 0.9) + 
    geom_node_text(aes(label = name), 
                   size = 3, color = col.bg, family = font.base) +
    theme_minimal(base_family = font.base) + 
    labs(title = "Gaussian DAG Structure for Synthesis", 
         subtitle = paste0("Nodes: ", as.character(q), 
                           ", Edges: ", as.character(DAG.edge.n), 
                           " (Theoretical Edge Inclusion Ratio: 0.2)"), 
         x = NULL, y = NULL) + 
    theme(
        title = element_text(color = col.tw, size = 10), 
        plot.subtitle = element_text(color = col.tw, size = 6), 
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.background = element_rect(fill = col.bg), 
    )
```
```{r}
set.seed(2022)
L2 <- L
L2[7, 2, 1] <- runif(1)
print(L2[7, 2, 1])

# Joint
causaleffect(targets = c(2, 7), response = 1, 
             L = L2[, , 1], D = D[, , 1])
# Single
causaleffect(targets = c(2), response = 1, 
             L = L2[, , 1], D = D[, , 1])
causaleffect(targets = c(7), response = 1, 
             L = L2[, , 1], D = D[, , 1])  # != 0
```
As it can be observed, this time a single intervention on X_7 produces a non-zero causal effect on X_1, while jointly intervening on X_2 and X_7 makes the total causal effect of X_7 on X1 zero. The effect of X_7 on X_1 was in fact mediated by X_2.

## Esitimate Causal Effects
The identification and estimation of causal effects requires the specification of a DAG. When the DAG is unknown, function `get_causaleffect()` can be used. 
It applies to objects of class `bcdag`; the latter corresponds to the output of learn_DAG() and consists of a sample of size S from the posterior of DAGs and DAG parameters. In addition get_causaleffect() takes as input a numerical vector representing the labels of the intervened nodes (the intervention target) and a numerical value indicating the response variable. **Output of the get_causaleffect() function** is a sample of size S from the posterior distribution of the causal effect coefficients associated with the intervention targets:
```{r}
effects_out <- get_causaleffect(out, targets = c(2, 7), response = 1)
head(effects_out)
```

Additionally, if BMA = TRUE, get_causaleffect() returns a Bayesian Model Average (BMA) estimate of the causal effect coefficients:
```{r}
## BMA estimate of causal effects
round(get_causaleffect(
    out, targets = c(2, 7), response = 1, BMA = TRUE
), 3)

## True causal effects
round(causaleffect(
    targets = c(2, 7), response = 1, L = L[, , 1], D = D[, , 1]
), 3)
```


# References
- Castelletti, F, Mascaro, A (2021). “Structural learning and estimation of joint causal effects among network-dependent variables”. Statistical Methods & Applications, 30(5), 1289-1314.

- Castelletti F, Mascaro A (2022). “BCDAG: An R package for Bayesian structural and Causal learning of Gaussian DAGs.” arXiv pre-print.

- Nandy, P, Maathuis, MH., Richardson, TS (2017). “Estimating the effect of joint interventions from observational data in sparse high-dimensional settings”. The Annals of Statistics, 45(2), 647-674.

- Pearl J (2000). Causality: Models, Reasoning, and Inference. Cambridge University Press, Cambridge. ISBN 0-521-77362-8.
