---
title: "kssa"
author: "Maxwell"
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 15
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

https://github.com/SteffenMoritz/kssa

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)

library(kssa)
library(imputeTS)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")
# icon.grob <- png_to_grob()

theme_update(panel.grid.minor = element_blank())
```


# Dataset {.tabset .tabset-fade .tabset-pills}
Create 20% missing completely at random data in tsAirgapComplete time series from imputeTS
```{r Dataset}
airgap_na <- missMethods::delete_MCAR(
    as.data.frame(tsAirgapComplete), 0.2
)
```

# Convert co2_na to time series object
```{r}
# 12 yrs of data are extended to 29 yrs of data by replication.
airgap_na_ts <- ts(
    data = airgap_na, start = c(1959, 1), end = c(1997, 12), frequency = 12
)
```


# Known Sub-Sequence Algorithm
Apply the kssa algorithm with 5 segments, 10 iterations, 20% of missing data, and compare among all available methods in the package.
Remember that percentmd must match with the real percentage of missing data in the input co2_na_ts time series

- "all": compare among all methods automatically - Default
- "auto.arima": State space representation of an ARIMA model
- "StructTS": State space representation of a structural model
- "seadec": Seasonal decomposition with Kalman smoothing
- "linear_i": Linear interpolation
- "spline_i": Spline interpolation
- "stine_i": Stineman interpolation
- "simple_ma": Simple moving average
- "linear_ma": Linear moving average
- "exponential_ma": Exponential moving average
- "locf": Last observation carried forward
- "stl": Seasonal and trend decomposition with Loess

```{r}
kssa.res <- kssa(
    x_ts = airgap_na_ts, 
    start_methods = "all", 
    actual_methods = "all", 
    segments = 5, 
    iterations = 10, 
    percentmd = 0.2, 
    seed = 2022
)
```

```{r}
cat(class(kssa.res))
cat(length(kssa.res))
cat(length(kssa.res[[1]]))
```


# kssa_plot
```{r}
p <- kssa_plot(kssa.res, type = "complete", metric = "mase")
```
```{r fig.height=15}
p + theme_minimal(base_family = font.base) + 
    coord_flip() + 
    scale_fill_carto_d(palette = "Safe") + 
    labs(x = "", y = "Mean Absolute Scaled Error", 
         title = "Known Sub-Sequence Algorithm (KSSA)", 
         subtitle = "tsAirgap dataset: monthly airline passengers") + 
    annotate(geom = "text", x = 1, y = 2.25, 
             label = "@Maxwell_110", color = col.tw) + 
    theme(
        text = element_text(family = font.base), 
        line = element_line(color = col.os, size = 0.5), 
        rect = element_rect(size = 0.5), 
        legend.position = c(0.6, 0.5), 
        legend.key.size = unit(0.05, "npc")
    )
ggsave("fig/kssa_01.jpg", width = 8, height = 8 * 1.5)
```


# References

Benavides IF, Santacruz M, Romero-Leiton JP, Barreto C, Selvaraj JJ. Assessing methods for multiple imputation of systematic missing data in marine fisheries time series with a new validation algorithm. Aquaculture and Fisheries 2022. https://doi.org/10.1016/j.aaf.2021.12.013.
