---
title: "Doing Meta-Analysis in R"
subtitle: "A Hands-on Guide"
author: ""
date: '`r Sys.Date()`'
# bibliography: references.bib
# link-citations: true
# zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 15
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)  # in tidyverse
library(lubridate)  # in tidyverse

# Visualisation
library(ggplot2)  # RColorBrewer::display.brewer.all()
library(ggalt)  # geom_encircle
library(ggimage)  # geom_subview
library(patchwork)
library(broman)  # plot_crayons()
library(ggsci)  # Journal palette
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(PrettyCols)  # view_all_palettes(colourblind_friendly = TRUE)
library(extrafont)  # fonttable(); "Candara"
library(latex2exp)  # example: latex2exp::TeX("Equation: $\\lambda$")

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)  # tbl_summary; tbl_regression; https://cran.r-project.org/web/packages/gtsummary/vignettes/tbl_regression.html


pacman::p_load(meta, metafor, esc, dmetar)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")

fn.prefix <- "doing-meta-analysis_"
```


# Effect Sizes {.tabset .tabset-fade .tabset-pills}
## (Standardized) Mean Difference
### Between-Group Mean Difference
$$
MD_{between} = \bar{x}_1 - \bar{x}_2 \\
SE_{MD_{between}} = s_{pooled} \; \sqrt{\frac{1}{n_1} + \frac{1}{n_2}} \\
s_{pooled} = \sqrt{\frac{S^2_{x_1} + S^2_{x_2}}{n_1 + n_2 - 2}} \\
$$

### Between-Group Standardized Mean Difference (Cohen's d)
$$
SMD_{between} = \frac{\bar{x}_1 - \bar{x}_2}{s_{pooled}} \\
SE_{SMD_{between}} = \sqrt{\frac{n_1 + n_2}{n_1 \; n_2} + \frac{SMD^2_{between}}{2 \; (n_1 + n_2)}}
$$

- Not necessary to do these calculations manually like we did here. For a meta-analysis of mean differences, we only have to prepare the following columns in our data set:
`n.e.` The number of observations in the intervention/experimental group.  
`mean.e.` The mean of the intervention/experimental group.  
`sd.e.` The standard deviation in the intervention/experimental group.  
`n.c.` The number of observations in the control group.  
`mean.c.` The mean of the control group.  
`sd.c.` The standard deviation in the control group.  

```{r}
# Define the data we need to calculate SMD/d
# This is just some example data that we made up
grp1m <- 50   # mean of group 1
grp2m <- 60   # mean of group 2
grp1sd <- 10  # sd of group 1
grp2sd <- 10  # sd of group 2
grp1n <- 100  # n of group1
grp2n <- 100  # n of group2

# Calculate effect size
esc_mean_sd(
    grp1m = grp1m,
    grp2m = grp2m,
    grp1sd = grp1sd,
    grp2sd = grp2sd,
    grp1n = grp1n,
    grp2n = grp2n
)
```


# Pooling Effect Sizes {.tabset .tabset-fade .tabset-pills}
## The Fixed-Effect Model

- The only reason why a study $k$'s observed effect size $\hat{\theta}_{k}$ deviates from $\theta$ is because of its sampling error $\epsilon_k$. $\hat{\theta}_{k} = \theta + \epsilon_k$
- All studies are estimators of the same true effect size. Yet, because every study can only draw somewhat bigger or smaller samples of the infinitely large study population, results are burdened by sampling error.
  
  
**Inverse-variance weighting**
$$
w_k = \frac{1}{s_k^2} \\
\hat{\theta} = \frac{\Sigma^K_{k=1} \; \hat{\theta}_k \; w_k}{\Sigma^K_{k=1} \; w_k}
$$
```{r}
data(SuicidePrevention)

# Calculate Hedges' g and the Standard Error
# - We save the study names in "study".
# - We use the pmap_dfr function to calculate the effect size for each row.
SP_calc <- pmap_dfr(SuicidePrevention, 
                    function(mean.e, sd.e, n.e, mean.c,
                             sd.c, n.c, author, ...){
                        esc_mean_sd(
                            grp1m = mean.e, grp1sd = sd.e, grp1n = n.e, 
                            grp2m = mean.c, grp2sd = sd.c, grp2n = n.c, 
                            study = author, 
                            es.type = "g"  # Hedges' g
                        ) %>% as.data.frame()
                    }) 

# The data set contains Hedges' g ("es") and standard error ("se")
glimpse(SP_calc)
```
```{r}
# Calculate the inverse variance-weights for each study
SP_calc$w <- 1 / SP_calc$se ^ 2

# Then, we use the weights to calculate the pooled effect
pooled_effect <- sum(SP_calc$w * SP_calc$es) / sum(SP_calc$w)
pooled_effect
```


# References
- https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/
