---
title: "Doing Meta-Analysis in R"
subtitle: "Section 5 - 9"
author: ""
date: '`r Sys.Date()`'
# bibliography: references.bib
# link-citations: true
# zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: yeti  # minty, flatly, litera, lumen, sandstone, spacelab, yeti
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 15
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)  # in tidyverse
library(lubridate)  # in tidyverse

# Visualisation
library(ggplot2)  # RColorBrewer::display.brewer.all()
library(ggalt)  # geom_encircle
library(ggimage)  # geom_subview
library(patchwork)
library(broman)  # plot_crayons()
library(ggsci)  # Journal palette
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(PrettyCols)  # view_all_palettes(colourblind_friendly = TRUE)
library(extrafont)  # fonttable(); "Candara"
library(latex2exp)  # example: latex2exp::TeX("Equation: $\\lambda$")

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)  # tbl_summary; tbl_regression; https://cran.r-project.org/web/packages/gtsummary/vignettes/tbl_regression.html


pacman::p_load(meta, metafor, esc, dmetar)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")

fn.prefix <- "doing-meta-analysis_"
```

# Between-Study Heterogeneity {.tabset .tabset-fade .tabset-pills}
```{r}

```


# Power Analysis {.tabset .tabset-fade .tabset-pills}
```{r}
power.analysis(
    d = 0.2,
    k = 10,
    n1 = 25,
    n2 = 25,
    p = 0.05
)
```
```{r}
power.analysis(
    d = 0.5,
    k = 10,
    n1 = 25,
    n2 = 25,
    p = 0.05,
    heterogeneity = "moderate"
)
```



# References
- https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/
- Bakbergenuly, Ilyas, David C. Hoaglin, and Elena Kulinskaya. 2020. “Methods for Estimating Between-Study Variance and Overall Effect in Meta-Analysis of Odds Ratios.” Research Synthesis Methods 11(3): 426–42. doi:10.1002/jrsm.1404.
