---
title: "oddnet"
subtitle: ""
author: "Maxwell"
date: '`r Sys.Date()`'
# bibliography: references.bib
# link-citations: true
# zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 15
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)  # RColorBrewer::display.brewer.all()
library(ggalt)  # geom_encircle
library(ggimage)  # geom_subview
library(patchwork)
library(broman)  # plot_crayons()
library(ggsci)  # Journal palette
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(PrettyCols)  # view_all_palettes(colourblind_friendly = TRUE)
library(extrafont)  # fonttable(); "Candara"

# Networks
library(ggraph)
# library(tidygraph)
library(oddnet)
library(igraph)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")

fn.prefix <- "oddnet_vignettes"

font.base <- "Georgia"
col.bg <- "#524A4E"  # Background
col.edge <- "#FDEFF4"  # Pale Pink
col.node <- "#4ECCA3"  # Teal
ggplot2::theme_set(theme_minimal(base_family = font.base))
```


# Dataset {.tabset .tabset-fade .tabset-pills}
Erdos-Renyi's random graph model:  
https://en.wikipedia.org/wiki/Erd%C5%91s%E2%80%93R%C3%A9nyi_model
```{r Dataset}
set.seed(2023)

networks <- list()
p.or.m.seq <- seq(from = 0.01, to = 0.1, length.out = 100)
p.or.m.seq[50] <- p.or.m.seq[50] + 0.2  # anomalous network

for(i in 1:100) {
    gr <- igraph::erdos.renyi.game(100, p.or.m = p.or.m.seq[i])
    networks[[i]] <- igraph::as_adjacency_matrix(gr)
}
```

# Plotting
## Networks 1
```{r}
network.1 <- networks[[1]]
gr.1 <- igraph::graph_from_adjacency_matrix(network.1)

p.1 <- ggraph(graph = gr.1, 
              layout = "igraph", algorithm = "nicely") + 
    geom_edge_link(
        # arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(3, "pt"), 
        end_cap = circle(3, "pt"), 
        color = col.edge, 
        edge_width = 0.25, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.node, size = 2, alpha = 0.9) + 
    theme_minimal(base_family = font.base) + 
    labs(title = "Erdos-Renyi Random Graph Network", 
         subtitle = "t = 1", 
         x = NULL, y = NULL) + 
    theme(
        title = element_text(color = col.tw, size = 10), 
        plot.subtitle = element_text(hjust = 0.1, size = 15), 
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.background = element_rect(fill = col.bg), 
    )
p.1
```

## Networks 30
```{r}
network.30 <- networks[[30]]
gr.30 <- igraph::graph_from_adjacency_matrix(network.30)

p.30 <- ggraph(graph = gr.30, 
               layout = "igraph", algorithm = "nicely") + 
    geom_edge_link(
        # arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(3, "pt"), 
        end_cap = circle(3, "pt"), 
        color = col.edge, 
        edge_width = 0.25, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.node, size = 2, alpha = 0.9) + 
    theme_minimal(base_family = font.base) + 
    labs(title = "", 
         subtitle = "t = 30", 
         x = NULL, y = NULL) + 
    theme(
        title = element_text(color = col.tw, size = 15), 
        plot.subtitle = element_text(hjust = 0.1, size = 15), 
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.background = element_rect(fill = col.bg), 
    )
p.30
```


## Networks 60
The networks have increased in connectivity because the edge probability increases over time. Let’s look at the anomalous network at t = 50.
```{r}
network.60 <- networks[[60]]
gr.60 <- igraph::graph_from_adjacency_matrix(network.60)

p.60 <- ggraph(graph = gr.60, 
               layout = "igraph", algorithm = "nicely") + 
    geom_edge_link(
        # arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(3, "pt"), 
        end_cap = circle(3, "pt"), 
        color = col.edge, 
        edge_width = 0.25, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.node, size = 2, alpha = 0.9) + 
    theme_minimal(base_family = font.base) + 
    labs(title = "", 
         subtitle = "t = 60 (> 50)", 
         x = NULL, y = NULL) + 
    theme(
        title = element_text(color = col.tw, size = 15), 
        plot.subtitle = element_text(hjust = 0.1, size = 15), 
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.background = element_rect(fill = col.bg), 
    )
p.60
```


# Features
```{r}
avg_deg <- rep(0, 100)
max_deg <- rep(0, 100)
for(i in 1:100){
    network <- networks[[i]]
    gr <- igraph::graph_from_adjacency_matrix(network)
    avg_deg[i] <- mean(degree(gr))
    max_deg[i] <- max(degree(gr))
}
```

```{r}
p.f <- ggplot(data = tibble(t = seq(100), avg.deg = avg_deg)) + 
    geom_point(aes(x = t, y = avg_deg), col = col.node) + 
    geom_line(aes(x = t, y = avg_deg), col = col.node) + 
    labs(
        title = "A feature of Erdos-Renyi Random Graph", 
        y = "Average degree \nof the nodes"
    ) + 
    theme(plot.title = element_text(hjust = 0.05, vjust = -10))
p.f
```


# Find Anomalies
```{r}
# anomalous_networks(networks)
```


# Save Figure
```{r}
p.f + {p.1 + p.30 + p.60} + plot_layout(ncol = 1, heights = c(1, 2))

ggsave(paste0("fig/", fn.prefix, ".jpg"), 
       dpi = 200, width = 10, height = 6)
```


# References
- https://cran.r-project.org/web/packages/oddnet/vignettes/oddnet.html

