---
title: "ggdensity: vignettes"
subtitle: "https://github.com/jamesotto852/ggdensity"
author: "Maxwell"
date: '`r Sys.Date()`'
# bibliography: references.bib
# link-citations: true
# zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 10
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

***

ggdensity extends ggplot2 providing more interpretable visualizations of density estimates based on highest density regions (HDRs). ggdensity offers drop-in replacements for ggplot2 functions:

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggdensity)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)

# GeoSpatial Analysis
library(sf)  # simple feature
library(spdep)  # helpful functions for mapping
library(rmapshaper)  # helpful functions for mapping
library(tmap)  # map viewer
library(leaflet)  # map viewer
library(mapview)  # map viewer
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")
# icon.grob <- png_to_grob()

theme_update(panel.grid.minor = element_blank())
```

# Dataset
```{r}
df <- data.frame("x" = rnorm(1000), "y" = rnorm(1000))
```

- instead of ggplot2::geom_density_2d_filled(), use ggdensity::geom_hdr()
- instead of ggplot2::geom_density_2d(), use ggdensity::geom_hdr_lines()

# geom_density_2d_filled() vs. geom_hdr()
https://github.com/jamesotto852/ggdensity#geom_density_2d_filled-vsgeom_hdr

While it’s a nice looking plot, it isn’t immediately clear how we should understand it. That’s because geom_density_2d_filled() generates its contours as equidistant level sets of the estimated bivariate density, i.e. taking horizontal slices of the 3d surface at equally-spaced heights, and projecting the intersections down into the plane. So you get a general feel of where the density is high, but not much else. To interpret a contour, you would need to multiply its height by the area it bounds, which of course is very challenging to do by just looking at it.

geom_hdr() tries to get around this problem by presenting you with regions of the estimated distribution that are immediately interpretable:

```{r fig.height=20}
p <- ggplot(df, aes(x, y)) + coord_equal()
p1 <- p + geom_density_2d_filled(contour_var = "ndensity")
p2 <- p + geom_hdr(probs = seq(0, 0.9, 0.1))

p1 / p2
```


# References
