---
title: "CLTVTools Vignette 01"
subtitle: ""
author: "Maxwell"
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
zotero: "My Library"
# abstract: \singlespacing Abstract which has to be long enough to 
#   take multiple lines otherwise one does not see the effect of single-spacing.
output: 
    html_document:
        number_sections: TRUE
        fig_caption: TRUE
        toc: TRUE
        toc_depth: 5
        toc_float: TRUE
        theme:
            bootswatch: minty
            # bootswatch: cosmo
        highlight: espresso  # espresso, tango, zenburn
        code_folding: show
        # code_download: TRUE
        fig_width: 15
editor_options: 
    chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

***

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)

library(CLVTools)
```

## Environment
```{r Environment, echo=FALSE, message=FALSE}
source(file = "utility/environments.R")
source(file = "utility/helper_functions.R")

fn.prefix <- "fig/CLVTools_"
# icon.grob <- png_to_grob()
```


# Dataset {.tabset .tabset-fade .tabset-pills}
```{r Dataset}
data("apparelTrans")
glimpse(apparelTrans)
```


# Initialize the CLV-Object
https://www.clvtools.com/articles/CLVTools.html#clvdata
```{r}
clv.apparel <- clvdata(apparelTrans,  
                       date.format="ymd", 
                       time.unit = "week",
                       estimation.split = 40,
                       name.id = "Id",
                       name.date = "Date",
                       name.price = "Price")
summary(clv.apparel)
```


# Estimate Model Parameters
https://www.clvtools.com/articles/CLVTools.html#estimate  
Fits Pareto/NBD models on transactional data with and without covariates.
```{r}
est.pnbd <- pnbd(clv.data = clv.apparel)

est.pnbd <- pnbd(
    clv.data = clv.apparel, 
    start.params.model = c(r = 1, alpha = 2, s = 1, beta = 2), 
    optimx.args = list(control = list(trace = 5), method = "Nelder-Mead" )
)

est.pnbd
```

```{r}
#Full detailed summary of the parameter estimates
summary(est.pnbd)

#Extract the coefficients only
coef(est.pnbd)

#Extract the confidence intervals
confint(est.pnbd)
```

```{r}
# LogLikelihood at maximum
logLik(est.pnbd)

# Variance-Covariance Matrix at maximum
vcov(est.pnbd)
```


# Predict Customer Behavior
https://www.clvtools.com/articles/CLVTools.html#predict
```{r}
results <- predict(est.pnbd)
print(results)
```
```{r}
predict(est.pnbd, prediction.end = 30)
```
```{r}
predict(est.pnbd, prediction.end = "2006-05-08")
```

# Plotting
https://www.clvtools.com/articles/CLVTools.html#plotting
## Weekly Tracking
```{r}
p1 <- plot(clv.apparel)
p1 <- p1 + theme(
    text = element_text(face = font.face), 
    plot.title = element_text(family = font.base, face = font.face),
    axis.title = element_text(family = font.base, face = font.face)
)
p1
```

## Density of Customer's Mean Time betwween Transactions
```{r}
p2 <- plot(clv.apparel, which="interpurchasetime")
p2 <- p2 + theme(
    text = element_text(face = font.face), 
    plot.title = element_text(family = font.base, face = font.face), 
    axis.title = element_text(family = font.base, face = font.face)
)
p2
```

## Weekly Tracking with Prediction
```{r}
p3 <- plot(est.pnbd)
p3 <- p3 + 
    labs(subtitle = "Fitting: 2005-01 to 2005-10", 
         caption = "@Maxwell_110") + 
    theme(
    text = element_text(face = font.face), 
    plot.title = element_text(family = font.base, face = font.face, size = 9), 
    plot.subtitle = element_text(size = 7), 
    plot.caption = element_text(size = 4, color = "gray70"), 
    axis.title = element_text(family = font.base, face = font.face, size = 7), 
    legend.title = element_blank(), 
    legend.text = element_text(size = 5), 
    legend.position = c(0.8, 0.9)
)
p3
ggsave(paste0(fn.prefix, "tracking_w_pred.jpg"), width = 5, height = 3.5)
```


# References


