---
title: "Survival Analysis"
subtitle: "Cox proportional hazard model"
author: "HoxoMaxwell"
date: '`r Sys.Date()`'
output: 
   html_document:
      number_sections: true
      toc: true
      toc_depth: 5
      toc_float: true
      theme: cosmo
      highlight: tango
      code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Load Library
survminer cheatsheet  
https://rpkgs.datanovia.com/survminer/survminer_cheatsheet.pdf
```{r}
library(survival)
library(survminer)
library(MASS)
library(DT)
```

# Load Data
```{r}
data("kidney")
datatable(
  kidney,
  options = list(pageLength = 10, 
                 autoWidth = TRUE,
                 scrollX = TRUE,
                 columnDefs = list(
                   list(width = '75px',
                        targets = seq(1, dim(kidney)[2])))),
  class = 'cell-border stripe'
  )
```

- patient:	 id
- time:	 time
- status:	 event status
- age:	 in years
- sex:	 1=male, 2=female
- disease:	 disease type (0=GN, 1=AN, 2=PKD, 3=Other)
- frail:	 frailty estimate from original paper

# Cox Proportional Hazard Analysis
## Build Survival Time Data Object
```{r}
Surv(kidney$time, kidney$status)
```

## Fit
```{r}
kid.cox <- coxph(Surv(time = time, time2 = status) ~ sex + disease + age,
                 data = kidney)
kid.cox
```
## Summary
```{r}
summary(kid.cox)
ggforest(model = kid.cox, data = kidney) 
```

## Plot Survival Curve
```{r, fig.width=10, fig.height=8}
kid.fit <- survfit(kid.cox)
ggsurvplot(
    fit = kid.fit,
    data = gehan,
    conf.int = T,
    pval = F,
    risk.table = F,
    cumevents = F,
    cumcensor = F,
    tables.height = 0.15
)
```

## Residuals
```{r}
resid(kid.cox)
```

ref:  
https://rpkgs.datanovia.com/survminer/reference/ggcoxdiagnostics.html
```{r}
ggcoxdiagnostics(
    fit = kid.cox,
    type = "martingale",
    ox.scale = "observation.id",
    hline = T, sline = T, sline.se = T, 
    hline.col = "slategray", hline.size = 1, hline.alpha = 0.7, 
    sline.col = "deeppink", sline.size = 2, sline.alpha = 0.2, 
    point.col = "dimgray", point.size = 1.5, point.alpha = 1.0,
    title = "Residual Analysis", 
    subtitle = "Data: Kidney",
    caption = "HoxoMaxwell",
    ggtheme = theme_light()
    )
```

```{r, fig.width=10, fig.height=10}
ggcoxdiagnostics(
    fit = kid.cox,
    type = "schoenfeld",
    ox.scale = "observation.id",
    hline = T, sline = T, sline.se = T, 
    hline.col = "slategray", hline.size = 1, hline.alpha = 0.7, 
    sline.col = "deeppink", sline.size = 2, sline.alpha = 0.2, 
    point.col = "dimgray", point.size = 1.5, point.alpha = 1.0,
    title = "Residual Analysis", 
    subtitle = "Data: Kidney",
    caption = "HoxoMaxwell",
    ggtheme = theme_light()
    )
```

## Proportional Haxard Test
```{r}
kid.ptest <- cox.zph(kid.cox)
```
```{r, fig.width=10, fig.height=10}
ggcoxzph(
    fit = kid.ptest, 
    resid = T,
    se = T,
    df = 4,
    nsmo = 40,
    # var = ,
    point.col = "dimgray", point.size = 1.5, point.alpha = 1.0,
    caption = "HoxoMaxwell",
    ggtheme = theme_light()
    )
```

## Adjusted Survival Curves
```{r}
ggadjustedcurves(kid.cox, variable = "sex", data = kidney)
```
```{r}
kidney[["cat_age"]] <- cut(kidney$age, seq(0, 80, 20))
ggadjustedcurves(kid.cox, variable = "cat_age", data = kidney)
```

## Interaction Terms
### All terms
```{r}
kid.cox2 <- coxph(Surv(time = time, time2 = status) ~ (sex + disease) ^ 2, 
                 data = kidney)
kid.cox2
```

### Variable Selection with stepAIC
```{r}
stepAIC(kid.cox2)
```

