---
title: "MCMC scheme for posterior inference of Gaussian DAG models: the learn_DAG() function"
subtitle: "R package: BCDAG vignettes vol.2"
author: ""
date: '`r Sys.Date()`'
# bibliography: references.bib
link-citations: true
abstract: \singlespacing This is the second of a series of three vignettes for the R package BCDAG. In this vignette we focus on function `learn_DAG()`, which implements a Markov Chain Monte Carlo (MCMC) algorithm to sample from the joint posterior of DAG structures and DAG-parameters under the Gaussian assumption.
output: 
   html_document:
      number_sections: TRUE
      fig_caption: TRUE
      toc: TRUE
      toc_depth: 5
      toc_float: TRUE
      theme:
         bootswatch: minty
         # bootswatch: cosmo
      highlight: espresso
      # espresso, tango, zenburn
      code_folding: show
      # code_download: TRUE
      fig_width: 10
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# knitr::opts_knit$set(root.dir = ".")  # modify with the location of code
```

<br>

<br>

**MCMC scheme for posterior inference of Gaussian DAG models: the learn_DAG() function**  
https://cran.r-project.org/web/packages/BCDAG/vignettes/bcdag_learnDAG.html

# Preparations {.tabset .tabset-fade .tabset-pills}
## Packages
```{r Packages, echo=FALSE, message=FALSE}
# Data manipulation
library(tidyverse)
library(data.table)
library(janitor)  # janitor::row_to_names()
library(DT)
library(stringr)
library(lubridate)

# Visualisation
library(ggplot2)
library(ggalt)  # geom_encircle
library(patchwork)
library(broman)  # plot_crayons()
library(rcartocolor)  # display_carto_all(); https://bit.ly/3Itq5kB
library(extrafont)  # "Candara"

# Interactivity
library(plotly)
library(crosstalk)

# Descriptive Statistics
library(mice)
library(ggmice)  # see src/imputation/ggmice_XX.Rmd
library(psych)
library(gtsummary)

# Parallel
library(purrr)
library(foreach)
library(doParallel)

# BCDAG
library(ggraph)
library(tidygraph)
library(BCDAG)

source(file = "utility/helper_functions.R")
```

## Environment
```{r Environment}
# rgb(73, 81, 112, maxColorValue = 255)
col.tw <- "#dbd7d2"  # Timberwolf
col.os <- "#414a4c"  # Outer Space
col.rp <- "#7851a9"  # Royal Purple
col.mb <- "#1a4876"  # Midnight Blue
col.pb <- "#1ca9c9"  # Pacific Blue
col.cb <- "#b0b7c6"  # Cadet Blue
col.el <- "#ceff1d"  # Electric Lime
col.cg <- "#00cc99"  # Caribbean Green
col.mt <- "#ff8243"  # Mango Tango
col.rm <- "#e3256b"  # Razzmatazz
col.sl <- "#fc2847"  # Scarlet

col.bg <- "#524A4E"  # Background
col.edge <- "#FDEFF4"  # Pale Pink
col.node <- "#4ECCA3"  # Teal

col.plos.yellow <- "#D6E13D"  # PLOS ONE Yellow
col.plos.pink <- "#CF00A3"  # PLOS ONE Pink
col.bmc.gray <- "#FFFDF6"  # BMC Gray
col.bmc.purple <- "#BF12F8"  # BMC Purple
col.bmc.navy <- "#1A2E4F"  # BMC Navy Blue
col.bmc.blue <- "#014A81"  # BMC Blue
col.bmc.sky <- "#0F99BE"  # BMC Sky Blue
col.bmc.green <- "#004940"  # BMC Deep Green
col.bmc.pink <- "#EF255F"  # BMC Pink

font.base <- "Candara"
# font.base <- "Times New Roman"
theme_set(theme_minimal(base_family = font.base))
options(dplyr.summarise.inform = TRUE)
# Twitter Logo and Acc
# Usage: add a code line below
# annotation_custom(grob = icon.grob, xmin = 1, xmax = 1, ymin = 1, ymax = 1)
icon.arr <- png::readPNG("fig/twitter_green.png")
icon.grob <- grid::rasterGrob(icon.arr, 
                              width = unit(.55 * 1., "cm"), 
                              height = unit(0.44 * 1., "cm"), 
                              interpolate = FALSE)
```

# Generating Data
## Generating a binary adjacency matrix w/ `rDAG` function
```{r}
set.seed(2022)

q <- 20
w <- 0.5
DAG <- rDAG(q = q, w = w)
class(DAG)
DAG
```

## Generating Gaussian DAG parameters w/ `rDAGWishart` function
```{r}
a <- q
U <- diag(1, q)
outDL <- rDAGWishart(n = 2, DAG, a, U)
class(outDL)

L <- outDL$L
class(L)
dim(L)
class(L[, , 1])
diag(L[, , 1])

D <- outDL$D
class(D)
dim(D)
class(D[, , 1])
sum(D[, , 1] > 0)
```

# Generating Data from a Gaussian DAG

<br>

$p(X_1, ..., X_q, L, D, \mathfrak{D}) = p(X_1, ..., X_q | L, D, \mathfrak{D}) \cdot p(L, D | \mathfrak{D}) \cdot p(\mathfrak{D})$

$p(X_1, ..., X_q | L, D, \mathfrak{D}) = N_{q} (0, \; (L D^{-1} L^T)^{-1})$  

$p(L, D | \mathfrak{D}) = \mathrm{DAG \text{-} Wishart} (a^{\mathfrak{D}}_c, U)$

$p(\mathfrak{D}) = \omega^{|S_{\mathfrak{D}}|} \; (1 - \omega)^{\frac{q(q-1)}{2} - |S_{\mathfrak{D}}|}$

$L$ and $D$: model parameters providing the decomposition of the precision matrix  
$U$ and $a^{\mathfrak{D}}_c$: hyper-parameters of prior to $(L, D)$  
$\omega$ and $S_{\mathfrak{D}}$: hyper-parameters of prior to $\mathfrak{D}$;  a Binomial distribution on the number of edges in the graph

```{r}
# Precision matrix: L D^-1 L^T
Omega <- L[, , 1] %*% solve(D[, , 1]) %*% t(L[, , 1])

# Covariance matrix
Sigma <- solve(Omega)

# MultiVariate Normal Distribution
x.mvn <- mvtnorm::rmvnorm(n = 10, sigma = Sigma)
dim(x.mvn)
```


# Visualization of generated DAG via `tidygraph` and `ggraph`
Reference: https://ggplot2-book.org/networks.html#networks  

```{r fig.height=10}
dag.tg <- tidygraph::as_tbl_graph(DAG, directed = TRUE)
class(dag.tg)
dag.tg

set_graph_style(plot_margin = margin(1, 1, 1, 1))
class(ggraph(dag.tg))

p1 <- ggraph(dag.tg, layout = "circlepack") + 
    geom_edge_link(
        arrow = arrow(angle = 15, length = unit(3, "pt")), 
        start_cap = circle(3, "pt"), 
        end_cap = circle(3, "pt"), 
        color = col.edge, 
        edge_width = 0.25, edge_alpha = 0.75
    ) + 
    geom_node_point(color = col.node, size = 2, alpha = 0.9) + 
    theme_minimal(base_family = font.base) + 
    labs(title = "Gaussian DAG Structure for Synthesis", 
         subtitle = "Nodes = 20, Edge Inclusion Ratio = 0.5", 
         x = NULL, y = NULL) + 
    theme(
        title = element_text(color = col.tw, size = 15), 
        axis.text = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        plot.background = element_rect(fill = col.bg), 
    ) + 
    annotation_custom(grob = icon.grob, 
                      xmin = 4.3, xmax = 4.3, 
                      ymin = 9.5, ymax = 9.5) + 
    annotate(geom = "text", x = 6.75, y = 9.5, 
             label = "@Maxwell_110", alpha = 0.3, 
             size = 4, family = "Candara", color = col.node)
plot(p1)
ggsave("fig/BCDAG-02-1.png", dpi = 300, width = 5, height = 5)
```


# MCMC Sampler for DAG Parameters
## learn_DAG
Function `learn_DAG()` implements an MCMC algorithm to sample from the joint posterior of DAGs and DAG parameters. 
This is based on a Partial Analytic Structure (PAS) algorithm (Godsill, 2012) which, at each iteration:

1. Updates the DAG through a Metropolis-Hastings (MH) step where, given the current DAG, a new (direct successor) DAG is drawn from a suitable proposal distribution and accepted with a probability given by the MH acceptance rate (see also section A note on fast = TRUE).
2. Samples from the posterior distribution of the (updated DAG) parameters; see also Castelletti & Consonni (2021) for more details.

NOTE:  
`> there is no package called graph`  
See below solution,  
https://stackoverflow.com/questions/31252102/bayesian-network-with-r  

> install.packages("BiocManager")  
> BiocManager::install(c("graph", "Rgraphviz"))  

```{r}
out <- learn_DAG(
    S = 5000, 
    burn = 1000, 
    data = x.mvn, 
    # In this tutorial, these hyperparameter are known.
    a = a, U = U, w = w, 
    fast = FALSE, save.memory = FALSE, collapse = FALSE, verbose = FALSE
)
```

## Outputs
```{r}
class(out)
names(out)
glimpse(out)
```

```{r, fig.height=4, fig.width=8}
dim(out$Graphs)
data.frame(
    sample.no = 1:5000, 
    inclusion.ratio = colSums(out$Graphs, dims = 2) / (q * (q - 1) / 2 + q)
) %>% ggplot() + 
    geom_point(aes(x = sample.no, y = inclusion.ratio), 
               color = col.node, size = 0.5) + 
    lims(y = c(0, 1)) + 
    labs(title = "MCMC Sampling for Edge Inclusion", 
         subtitle = "Edge Inclusion Ratio = 0.5 (Theoretical)", 
         x = "Sample No", y = "Edge Inclusion Ratio") + 
    theme_minimal(base_family = font.base) + 
    theme(
        title = element_text(color = col.node, size = 15), 
        axis.title = element_text(color = col.node, size = 12), 
        axis.text = element_text(color = col.node, size = 10), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_line(color = col.edge, linetype = 3,
                                        size = 0.1),
        plot.background = element_rect(fill = col.bg)
    ) + 
    annotation_custom(grob = icon.grob, 
                      xmin = 4100, xmax = 4100, 
                      ymin = 0.9, ymax = 0.9) + 
    annotate(geom = "text", x = 4550, y = 0.9, 
             label = "@Maxwell_110",  alpha = 0.3, 
             size = 4, family = "Candara", color = col.node)
ggsave("fig/BCDAG-02-2.png", dpi = 300, width = 8, height = 4)
```

